---
output: 
  word_document:
    reference_docx: ../report_template_v1.5.dotx
---

```{r setup, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}

# =======================================================================================
# ARU assessment data and diagnostics
# 
# 21/04/2020 First coding; adapted from WGWIDE code
# 25/04/2021 Fully adapted to inclusion of plots for WGDEEP report
# =======================================================================================

# devtools::install_github("fishfollower/SAM/stockassessment")

# try(if(packageVersion("stockassessment")=="0.5.4") stop("wrong version of Stockassessment package"))

require("knitr")
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 

rm(list=ls())

# library(devtools)
library(tidyverse)    # tidying packages
library(readxl)       # read excel
library(scales)
library(RColorBrewer)
library(viridis)
library(pander)
library(stockassessment)
library(ggrepel)

# devtools::install_github("einarhjorleifsson/readsam")
library(readsam)

# lowcase function
source("lowcase.r")
source("theme_publication.r")
source("sam_process_error.r")

datapath <- "C:/DATA/Onedrive - PFA/Documents/iWGDEEP/2021/06. Data/aru.27.5b6a"

# Function to download fit from web to local computer
getfitfromweb <- function(nam){
  dir.create(file.path(nam))
  download.file(file.path("https://www.stockassessment.org/datadisk/stockassessment/userdirs/user233",
                          nam, "run","model.RData"), file.path(nam, "model.RData"))
  load(file.path(nam, "model.RData"))
  return(fit)
  }
# getfitfromweb(nam="WGDEEP_2020")


getparsfromfit <- function(fit) {
  # a  <- gsub("fit_", "", fit) 
  df <- 
    data.frame(stockassessment::partable(fit)) %>% 
    add_rownames() %>%  
    separate(rowname, into=c("parameter", "parnumber"), sep="_")
  return(df)
}

nam         <- "ARU.27.5b6a_WGDEEP_2021"
assessments <- c("ARU.27.5b6a_WGDEEP_2020_", "ARU.27.5b6a_WGDEEP_2021", "Aru.27.5b6a_wgdeep2021_trial_deepsurv")
user        <- "user233"
web         <- TRUE

rp          <- data.frame(var      = c("fbar","ssb","ssb"),
                          refpoint = c("fmsy","blim","msybtrigger"),
                          value    = c(0.2, 59730, 82999))

# read input, results and parameter data
ibya <- rbya <- rby <- pars <- pe_rby <- pe_rbya <- data.frame()
for (a in assessments) {
  # print(a)
  assign(paste0("fit_", a), getfitfromweb(a))  
  assign(paste0("ibya_",a), mutate(readsam::read_ibya(a, web = web, user = user), file=a))  
  assign(paste0("rbya_",a), mutate(readsam::read_rbya_sam(get(paste0("fit_", a)), get(paste0("ibya_",a))), file=a) )  
  assign(paste0("rby_",a) , mutate(readsam::read_rby_sam(get(paste0("fit_", a)), get(paste0("ibya_",a))), file=a) )
  assign(paste0("pars_",a), mutate(getparsfromfit(get(paste0("fit_", a))), file=a)) 
  t <- sam_process_error(get(paste0("rbya_",a)),plus_group = TRUE, plot_it=FALSE)
  assign(paste0("pe_rbya_",a), mutate(t[[1]], file=a)) 
  assign(paste0("pe_rby_",a), mutate(t[[2]], file=a)) 
  
  ibya <- bind_rows(ibya, get(paste0("ibya_",a)))
  rbya <- bind_rows(rbya, get(paste0("rbya_",a)))
  rby  <- bind_rows(rby, get(paste0("rby_",a)))
  pars <- bind_rows(pars, get(paste0("pars_",a)))
  pe_rbya <- bind_rows(pe_rbya, get(paste0("pe_rbya_",a)))
  pe_rby  <- bind_rows(pe_rby, get(paste0("pe_rby_",a)))
}

# identical(fit_ARU.27.5b6a_WGDEEP_2020_, fit_ARU.27.5b6a_WGDEEP_2021)





# Parameters
pars <- 
  pars %>% 
  mutate(parnumber = factor(parnumber, levels=unique(.$parnumber))) %>% 
  lowcase()

# url        <- paste("https://www.stockassessment.org/datadisk/stockassessment/userdirs",user,"/",nam,"/data/", sep="")  
# filestoget <- c("survey.dat")
# d          <- lapply(filestoget, function(f)download.file(paste(url,f,sep=""), f))

fleetnames <-
  bind_rows(
    data.frame(fleetname="catch", stringsAsFactors = FALSE),
    as.data.frame(names(FLCore::readFLIndices("survey.dat"))) %>% setNames("fleetname"),
    
  ) %>% 
  mutate(fleet = as.character(row_number())) %>% 
  mutate(fleetname  = gsub(" ","", fleetname)) %>% 
  mutate(fleetname2 = substr(fleetname, 1, 12))

fit  <- fit_ARU.27.5b6a_WGDEEP_2021
conf <- fit$conf
sd   <- fit$pl
data <- fit$data

ages   <- conf$minAge:conf$maxAge
fleets <- 1:nrow(conf$keyLogFsta)
years  <- data$years
nr     <- nrow(conf$keyLogFsta)
nc     <- ncol(conf$keyLogFsta)

t1 <- 
  as.data.frame(conf$keyLogFpar) %>% 
  rownames_to_column(var="fleet") %>% 
  pivot_longer(names_to="test", values_to="parnumber", cols=1:nc+1) %>% 
  bind_cols(age=rep(ages, nr)) %>% 
  filter(parnumber != -1) %>% 
  group_by(parnumber, fleet) %>% 
  summarise(ages = paste(age, collapse="_")) %>% 
  ungroup() %>% 
  distinct(parnumber, fleet, ages) %>% 
  mutate(partype = "LogFpar") %>% 
  left_join(fleetnames, by="fleet") %>% 
  mutate(parname = paste(fleetname2, ages))

t2 <- 
  as.data.frame(conf$keyVarObs) %>% 
  rownames_to_column(var="fleet") %>% 
  pivot_longer(names_to="test", values_to="parnumber", cols=1:nc+1) %>% 
  bind_cols(age=rep(ages, nr)) %>% 
  filter(parnumber != -1) %>% 
  group_by(parnumber, fleet) %>% 
  summarise(ages = paste(age, collapse="_")) %>% 
  ungroup() %>% 
  distinct(parnumber, fleet, ages) %>% 
  mutate(partype = "sdLogObs") %>% 
  left_join(fleetnames, by="fleet") %>% 
  mutate(parname = paste(fleetname2, ages))

t3 <- 
  as.data.frame(conf$keyVarObs) %>% 
  rownames_to_column(var="fleet") %>% 
  pivot_longer(names_to="test", values_to="parnumber", cols=1:nc+1) 

# "sdrep" parameter estimates TMB report
# "pl"    parameter list
# "plsd"  sd parameter estimates
# "data"  data
# "conf"  configuration
# "opt"   optimization
# "obj"   likelihood object
# "rep"   additional report
# "low"   limit for parameters
# "hig"   limit for parameters

# Generating a data frame with names in the fit list
l1 <- l2 <- l3 <- data.frame(stringsAsFactors = FALSE)
count <- 0
l1 <- as.data.frame(names(fit)) 
for (i in 1:nrow(l1)) {
  l2 <- bind_rows(l2, data.frame(level2 = names(fit[[i]]), level1 = l1[i,1]))
  for (j in 1:length(fit[[i]])) {
    count <- count + 1
    # print(paste(i,j,count,sep=" / "))
    if(!is.null(names(fit[[i]][[j]]))) {
      # print(unique(names(fit[[i]][[j]])))
      l3 <- bind_rows(l3, data.frame(level3 = unique(names(fit[[i]][[j]])), level2 = l2[count,1], level1 = l1[i,1]))
    } else {
      # print(unique(names(fit[[i]][[j]])))
      l3 <- bind_rows(l3, data.frame(level2 = l2[count,1], level1 = l1[i,1]))
    }
  }
}

l3 <- l3 %>% dplyr::select(level1, level2, level3)



```

**Handling of catch data for Greater argentine in 5b and 6a**

Martin Pastoors, `r format(Sys.time(), '%d/%m/%Y')`

Comparing catch in tonnes (from InterCatch, all ages included) with the estimated/calculated catch in tonnes from SAM assessment. 

```{r, echo=FALSE, fig.asp=0.8, fig.align="center", message=FALSE, warning=FALSE}

# observed catch in tonnes
ct <- 
  read.table("ct.dat", skip=5) %>% 
  as.data.frame() %>%
  setNames("estimate") %>% 
  mutate(
    estimate   = as.numeric(estimate),
    source="intercatch",
    var   = "catch", 
  ) %>% 
  bind_cols(year=fit$data$years) %>% 
  mutate(year  = as.numeric(year))

# assessment
ass <-
  rby %>% 
    filter(file %in% c("ARU.27.5b6a_WGDEEP_2021")) %>% 
    filter(variable %in% c("ssb","fbar","catch")) %>%
    # filter(variable %in% c("catch")) %>% 
    lowcase() %>% 
    rename(var=variable) %>%
    mutate(source="assessment") %>%
    mutate(year = ifelse(var=="catch", year+1994,year)) 

# plot comparing obs vs pred  
ass %>% 
  filter(var=="catch") %>% 
  ggplot(aes(x=year,y=estimate)) +
  theme_publication() +
  theme(plot.margin=grid::unit(c(0,0,0,0), "mm")) +
  theme(legend.position="none") +
  # assessment
  geom_line(aes(colour=source)) +
  geom_ribbon(aes(ymin=low, ymax=high, fill=source), alpha=0.3) +
  
  # observations
  geom_point(data=ct, aes(x=year,y=estimate), inherit.aes = FALSE) +

  expand_limits(y=0) +
  labs(x="", y="", title="catch (obs vs predicted") 


# plot showing the relative difference between obs vs predicted
ass %>% 
  bind_rows(ct) %>% 
  filter(var=="catch") %>% 
  dplyr::select(estimate, year, var, source) %>% 
  pivot_wider(names_from = source, values_from=estimate) %>% 
  mutate(rel_diff = 100*(assessment/intercatch - 1)) %>% 
  
  ggplot(aes(x=year,y=rel_diff)) +
  theme_publication() +
  theme(legend.position="none") +
  geom_bar(aes(fill=var), stat="identity", size=1) +
  labs(x="") 



```

##### page break

InterCatch catch in tonnes comparison: differentiating between catch at age 5 and higher and catch at age 4 and lower. In 2020, 10% of the caton was in the ages 4 and below. That means that the caton used in the SAM model was actually 15475 instead of 17149 t.

```{r, echo=FALSE, fig.asp=0.4, fig.align="center", message=FALSE, warning=FALSE}

# InterCatch

ic_cn <-
  readxl::read_xlsx(file.path(datapath, "20210426 InterCatch all data.xlsx"),
                    sheet="Canum",
                    skip=61,
                    col_names=TRUE) %>% 
  pivot_longer(names_to="age", values_to="canum", 2:ncol(.))

ic_cw <-
  readxl::read_xlsx(file.path(datapath, "20210426 InterCatch all data.xlsx"),
                    sheet="Weca",
                    skip=61,
                    col_names=TRUE) %>% 
  pivot_longer(names_to="age", values_to="weca", 2:ncol(.))

ic <-
  ic_cn %>% 
  left_join(ic_cw, by=c("year","age")) %>% 
  mutate(sop = canum * weca) %>% 
  mutate(age = as.integer(age)) %>% 
  mutate(inassessment = ifelse(age >= 5, TRUE, FALSE)) %>% 
  group_by(year, inassessment) %>% 
  summarise(sop = sum(sop, na.rm=TRUE))


ic %>% 
  ggplot(aes(x=year,y=sop)) +
  theme_publication() +
  geom_bar(aes(fill=inassessment), stat="identity", size=1) +
  labs(x="") 

ic %>% 
  ggplot(aes(x=year,y=sop)) +
  theme_publication() +
  geom_bar(aes(fill=inassessment), stat="identity", position="fill", size=1) +
  labs(x="") 

ic %>% 
  group_by(year) %>% 
  mutate(prop = 100* sop / sum(sop, na.rm=TRUE)) %>% 
  filter(inassessment == FALSE) %>% 
  
  ggplot(aes(x=year,y=prop)) +
  theme_publication() +
  geom_bar(stat="identity", size=1) +
  labs(x="", y="prop (%)", title = "Proportion caton not in assessment (age <= 4") 

ic %>% 
  group_by(year) %>%
  mutate(sop = as.integer(sop)) %>% 
  pivot_wider(names_from=inassessment, values_from=sop) %>% 
  setNames(c("year","age2_4","age5_21")) %>% 
  mutate(total = age2_4 + age5_21) %>% 
  mutate(prop = scales::percent( age2_4 / total)) %>% 
  pander::pandoc.table(style="simple", justify="right" )





```
##### page break

Comparison of summed logged observed catch numbers in InterCatch (age 2-21) and in SAM (age 5-21), and calculation of percentage difference (in SAM compared to InterCatch)

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}

# InterCatch
# Catch in numbers comparison

fleets <- unique(fit$data$aux[, "fleet"])
idx <- fit$data$aux[, "fleet"] %in% fleets


# log catch number at age
t <- 
  data.frame(
    age   = fit$data$aux[idx, "age"],
    fleet = as.character(fit$data$aux[idx, "fleet"]),
    year  = fit$data$aux[idx, "year"],
    obs   = fit$data$logobs[idx],
    pred  = fit$rep$predObs[idx]
  ) %>% 
  filter(fleet==1) %>% 
  mutate(res   = obs-pred) %>% 
  mutate(source="sam") %>% 
  
  bind_rows(
    ic_cn %>% 
      mutate(obs=log(canum)) %>% 
      mutate(source="ic") %>% 
      mutate(age=as.numeric(age))
  ) 

# plot obs in SAM and obs in IC
t %>%
  ggplot(aes(x=year, y=obs)) +
  theme_publication() +
  geom_point(aes(colour=source)) +
  geom_line(aes(colour=source)) +
  scale_x_continuous(breaks=scales::pretty_breaks(n=3)) +
  facet_wrap(~age, scales="free_y")


# plot obs in SAM and obs in IC, summed over ages
t %>%
  group_by(year, source) %>% 
  summarise(obs=sum(obs,na.rm=TRUE)) %>% 

  ggplot(aes(x=year, y=obs)) +
  theme_publication() +
  geom_point(aes(colour=source)) +
  geom_line(aes(colour=source)) +
  scale_x_continuous(breaks=scales::pretty_breaks(n=3)) 


t %>% 
  group_by(year, source) %>% 
  summarise(obs=sum(obs,na.rm=TRUE)) %>% 
  pivot_wider(names_from=source, values_from=obs) %>% 
  mutate(perc_diff = scales::percent(sam/ic-1)) %>% 
  pander::pandoc.table(style="simple", justify="right" )

```

##### page break

Catch in number comparison between predicted and observed (i.e. age 5 onwards) in SAM

```{r, echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}

# Catch in numbers comparison

fleets <- unique(fit$data$aux[, "fleet"])
idx <- fit$data$aux[, "fleet"] %in% fleets


# log catch number at age
t <- 
  data.frame(
    age   = fit$data$aux[idx, "age"],
    fleet = as.character(fit$data$aux[idx, "fleet"]),
    year  = fit$data$aux[idx, "year"],
    obs   = fit$data$logobs[idx],
    pred  = fit$rep$predObs[idx]
  ) %>% 
  filter(fleet==1) %>% 
  mutate(res   = obs-pred) 

# plot obs and pred
t %>%
  ggplot(aes(x=year, y=obs)) +
  theme_publication() +
  geom_point() +
  geom_line(aes(y=pred)) +
  scale_x_continuous(breaks=scales::pretty_breaks(n=3)) +
  facet_wrap(~age, scales="free_y")

# plot residuals by age
t %>%
  ggplot(aes(x=year, y=res)) +
  theme_publication() +
  geom_hline(yintercept=0) +
  geom_segment(aes(x=year, xend=year, yend=res, y=0)) +
  geom_point(fill="white", shape=21) +
  scale_y_continuous(limits = c(-max(abs(t$res), na.rm=TRUE), max(abs(t$res), na.rm=TRUE))) +
  scale_x_continuous(breaks=scales::pretty_breaks(n=3)) +
  facet_wrap(~age)


# plot residuals
t %>%
  group_by(year) %>% 
  summarise(
    obs = sum(obs, na.rm=TRUE),
    pred = sum(pred, na.rm=TRUE)
  ) %>% 
  mutate(res   = obs-pred) %>%
  mutate(resprop = res / obs) %>% 
  # View()

  ggplot(aes(x=year, y=res)) +
  theme_publication() +
  # geom_hline(yintercept=0) +
  geom_segment(aes(x=year, xend=year, yend=res, y=0)) +
  geom_point(fill="white", shape=21) +
  # scale_y_continuous(limits = c(-max(abs(t$res), na.rm=TRUE), max(abs(t$res), na.rm=TRUE))) +
  scale_x_continuous(breaks=scales::pretty_breaks(n=3)) 


```



```{r, echo=FALSE, fig.asp=0.8, fig.align="center", message=FALSE, warning=FALSE}



```