---
output: 
  word_document:
    reference_docx: ../report_template_v1.5.dotx
---

```{r setup, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}

# =======================================================================================
# ARU 5b6a assessment and forecast
# 
# 21/04/2020 First coding; adapted from WGWIDE code
# 25/04/2021 Fully adapted to inclusion of plots for WGDEEP report
# 30/04/2022 Adapted for WGDEEP 2022
# =======================================================================================

# devtools::install_github("fishfollower/SAM/stockassessment")

# try(if(packageVersion("stockassessment")=="0.5.4") stop("wrong version of Stockassessment package"))

require("knitr")
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 

options(dplyr.summarise.inform = FALSE)

rm(list=ls())

# library(devtools)
library(tidyverse)    # tidying packages
library(readxl)       # read excel
library(scales)
library(RColorBrewer)
library(viridis)
library(pander)
library(stockassessment)
# devtools::install_github("shfischer/FLfse/FLfse")
# library(FLfse)  # 
library(ggrepel)

# devtools::install_github("einarhjorleifsson/ggmisc")
# library(ggmisc)     # e.g. crayola

# devtools::install_github("einarhjorleifsson/readsam")
# library(readsam)

# lowcase function
source("lowcase.r")
source("theme_publication.r")
source("sam_process_error.r")
source("fix_forecast.r")
source("common.R")

#' Function for extracting an object from a .RData file created by R's save() command
#' Inputs: RData file, object name
extract_from_rdata <- function(file, object) {
  E <- new.env()
  load(file=file, envir=E)
  return(get(object, envir=E, inherits=F))
}

# Add units function
addUnits <- function(n) {
  labels <- dplyr::case_when(
    n < 1000 ~ as.character(n),  # less than thousands
    n < 1e6  ~ paste0(round(n/1e3), 'k'),  # in thousands
    n < 1e9  ~ paste0(round(n/1e6), 'M'), # in millions
    n < 1e12 ~ paste0(round(n/1e9), 'B'), # in billions
    n < 1e15 ~ paste0(round(n/1e12), 'T'), # in trillions
    TRUE     ~ 'too big!'
  )
  return(labels)
}


nam         <- "ARU.27.5b6a_WGDEEP_2022"
assessments <- c("ARU.27.5b6a_WGDEEP_2021", nam)
user        <- "user233"
web         <- TRUE
filepath    <- file.path("https://www.stockassessment.org/datadisk/stockassessment/userdirs",
                         user,
                         nam,
                         "run")
configpath    <- file.path("https://www.stockassessment.org/datadisk/stockassessment/userdirs",
                         user,
                         nam,
                         "conf")
localpath   <- file.path("C:/temp")

rp          <- data.frame(variable = c("fbar","ssb","ssb"),
                          refpoint = c("fmsy","blim","msybtrigger"),
                          value    = c(0.24, 59730, 82999)) %>% 
    mutate(variable = factor(variable, levels=c("ssb","fbar","rec")))


# a <- assessments[[2]]

# read input, results and parameter data

# ibya <- rbya <- rby <- obscatch <- pars <- opr <- resosa <- loo_rby <- retro_rby <- tibble()
# for (a in assessments) {
#   print(a)
# 
#   # set filepath
#   filepath    <- file.path("https://www.stockassessment.org/datadisk/stockassessment/userdirs",
#                            user,
#                            a,
#                            "run")
# 
#   # Load assessment RData files
#   # load(file=file.path(localpath, paste0(a, ".RData")))
#   
#   # fit
#   assign(paste0("fit_", a),
#          getfitfromweb(a))
# 
#   # flstock
#   assign(paste0("flstock_", a),
#          FLfse::SAM2FLStock(get(paste0("fit_", a))))
# 
#   #leaveout
#   fil <- file.path(filepath, "leaveout.RData")
#   if(RCurl::url.exists(fil)) {
#     # assign the leaveout.RData file
#     assign(paste0("loo_", a),
#            extract_from_rdata(url(fil), "LO"))
# 
#   } else {
#     # run the leave out analyse
#     assign(paste0("loo_", a),
#          stockassessment::leaveout(get(paste0("fit_", a))))
#   }
# 
#   #retro
#   fil <- file.path(filepath, "retro.RData")
#   if(RCurl::url.exists(fil)) {
#     # assign the RData file
#     assign(paste0("retro_", a),
#            extract_from_rdata(url(fil), "RETRO"))
# 
#   } else {
#     # run the analysis
#     assign(paste0("retro_", a),
#          stockassessment::retro(get(paste0("fit_", a)), year=5))
#   }
# 
#   # residuals                     # stockassessment:::procres(fit)
#   fil <- file.path(filepath, "residuals.RData")
#   if(RCurl::url.exists(fil)) {
#     # assign the RData file
#     assign(paste0("res_", a),
#            extract_from_rdata(url(fil), "RES"))
#     assign(paste0("procres_", a),
#            extract_from_rdata(url(fil), "RESP"))
# 
#   } else {
#     # run the analyses
#     assign(paste0("res_", a),
#          residuals(get(paste0("fit_", a))))
#     assign(paste0("procres_", a),
#          stockassessment::procres(get(paste0("fit_", a))))
#   }
# 
#   #ibya
#   assign(paste0("ibya_",a),
#          mutate(fishvice::sam_ibya(getfitfromweb(a)), file=a))
# 
#   #observed catch
#   assign(paste0("obscatch_",a),
#          get(paste0("ibya_",a)) %>%
#          mutate(est = oC * cW,
#                 variable="catch") %>%
#          group_by(year, variable, file) %>%
#          summarise(est = sum(est, na.rm=TRUE)))
# 
#   # rbya
#   assign(paste0("rbya_",a),
#          mutate(fishvice::sam_rbya(get(paste0("fit_", a))), file=a) )
# 
#   # rby
#   assign(paste0("rby_",a) ,
#          mutate(fishvice::sam_rby(get(paste0("fit_", a))), file=a) )
# 
#   # opr
#   assign(paste0("opr_",a) ,
#          mutate(fishvice:::sam_opr(get(paste0("fit_", a))), file=a) )
# 
#   # parameters
#   assign(paste0("pars_",a) ,
#          mutate(fishvice:::sam_partable(get(paste0("fit_", a))), file=a) )
# 
#   # resosa
#   assign(paste0("resosa_",a) ,
#          mutate(fishvice:::sam_one_observation_ahead_residuals(get(paste0("res_", a))), file=a) )
# 
#   # retro_rby
#   assign(paste0("retro_rby_", a),
#          get(paste0("rby_", a)) %>%  mutate(finalyear = max(get(paste0("rby_", a))$year)))
#   for (i in 1:length(get(paste0("retro_", a)))) {
#     assign(paste0("retro_rby_", a),
#        bind_rows(
#          get(paste0("retro_rby_", a)),
#          fishvice::sam_rby(get(paste0("retro_", a))[[i]]) %>%
#            mutate(finalyear = max(get(paste0("retro_", a))[[i]]$data$years))
#          )
#        )
#   }
# 
#   # loo_rby
#   assign(paste0("loo_rby_", a), tibble())
#   # i <- names(get(paste0("loo_", a)))[1]
#   for (i in names(get(paste0("loo_", a)))) {
#     assign(paste0("loo_rby_", a),
#        bind_rows(
#          get(paste0("loo_rby_", a)),
#          fishvice::sam_rby(get(paste0("loo_", a))[[i]]) %>%
#            mutate(leftout = i)
#          )
#        )
#   }
# 
#   # to do: process error
# 
#   # bind together
#   ibya     <- bind_rows(ibya, get(paste0("ibya_",a)))
#   rbya     <- bind_rows(rbya, get(paste0("rbya_",a)))
#   rby      <- bind_rows(rby, get(paste0("rby_",a)))
#   obscatch <- bind_rows(obscatch, get(paste0("obscatch_",a)))
#   pars     <- bind_rows(pars, get(paste0("pars_",a)))
#   opr      <- bind_rows(opr, get(paste0("opr_",a)))
#   resosa   <- bind_rows(opr, get(paste0("opr_",a)))
#   retro_rby  <- bind_rows(retro_rby, get(paste0("retro_rby_",a)) %>% mutate(file=a))
#   loo_rby  <- bind_rows(loo_rby, get(paste0("loo_rby_",a)) %>% mutate(file=a))
# 
#   save(
#     list=as.character(
#       c(
#         paste0("fit_",a),
#         paste0("flstock_",a),
#         paste0("loo_",a),
#         paste0("retro_",a),
#         paste0("res_",a),
#         paste0("procres_",a),
#         paste0("ibya_",a),
#         paste0("rbya_",a),
#         paste0("rby_",a),
#         paste0("obscatch_",a),
#         paste0("pars_",a),
#         paste0("opr_",a),
#         paste0("resosa_",a),
#         paste0("retro_rby_",a),
#         paste0("loo_rby_",a)
#     )),
#     file=file.path(localpath, paste0(a, ".RData"))
#   )
# 
# }


# Load assessment RData files
for (a in assessments) {
  load(
    file=file.path(localpath, paste0(a, ".RData"))
  )
}

# combine overall assessment results
ibya <- rbya <- rby <- obscatch <- pars <- opr <- resosa <- loo_rby <- retro_rby <- tibble()
for (a in assessments) {
  ibya     <- bind_rows(ibya, get(paste0("ibya_",a)))
  rbya     <- bind_rows(rbya, get(paste0("rbya_",a)))
  rby      <- bind_rows(rby, get(paste0("rby_",a)))
  obscatch <- bind_rows(obscatch, get(paste0("obscatch_",a)))
  pars     <- bind_rows(pars, get(paste0("pars_",a)))
  opr      <- bind_rows(opr, get(paste0("opr_",a)))
  resosa   <- bind_rows(opr, get(paste0("opr_",a)))
  retro_rby<- bind_rows(retro_rby, 
                        get(paste0("retro_rby_",a)) %>% mutate(file=a))
  loo_rby  <- bind_rows(loo_rby, 
                        get(paste0("loo_rby_",a)) %>% mutate(file=a))
}


```

**Overview of SAM assessments for Greater argentine in 5b and 6a**

**`r paste(assessments, collapse="; ")`**

Martin Pastoors, `r format(Sys.time(), '%d/%m/%Y')`

**Tables**

**Table 1. Summary of the assessment**

```{r, echo=FALSE, fig.asp=0.8, fig.align="center", message=FALSE, warning=FALSE}

fit <- get(paste("fit", assessments[2], sep="_"))

stamp<-gsub('-[[:digit:]]{4}$','',gsub(':','.',gsub(' ','-',gsub('^[[:alpha:]]{3} ','',date()))))

catch<-tsbtable(fit)
colnames(catch)<-c("CATCH","Low", "High")

tab.summary <- as.data.frame(cbind(summary(fit), catch)) %>% rownames_to_column(var="year")
pander::pandoc.table(tab.summary,
                     style="simple", split.tables=200, justify="right", missing=".")

```

\newpage

**Table 2. Estimated fishing mortality at age**

```{r, echo=FALSE, fig.asp=0.8, fig.align="center", message=FALSE, warning=FALSE}

pander::pandoc.table(
  as.data.frame(faytable(fit)) %>%  rownames_to_column(var="year"),
  style="simple", split.tables=200, justify="right", missing=".",
  round=3)

```

\newpage

**Table 3. Estimated stock numbers at age**

```{r, echo=FALSE, fig.asp=0.8, fig.align="center", message=FALSE, warning=FALSE}

pander::pandoc.table(
  as.data.frame(ntable(fit)) %>%  mutate(across(.cols = everything(), ~ ./1)) %>% rownames_to_column(var="year"),
  style="simple", split.tables=200, justify="right", missing=".",
  round=0)

```

\newpage

**Table 4. model parameters**

```{r, echo=FALSE, fig.asp=0.8, fig.align="center", message=FALSE, warning=FALSE}

pander::pandoc.table(
  as.data.frame(partable(fit)) %>%  rownames_to_column(var="name"),
  style="simple", split.tables=200, justify="right", missing=".")

```

**Table 5. model fitting**

```{r, echo=FALSE, fig.asp=0.8, fig.align="center", message=FALSE, warning=FALSE}

pander::pandoc.table(
  as.data.frame(modeltable(fit)) %>%  rownames_to_column(var="model"),
  style="simple", split.tables=200, justify="right", missing=".",
  round=3,
  caption="Table 5. Model fitting")

```

**Table 6. Selected standard deviations**

```{r, echo=FALSE, fig.asp=0.8, fig.align="center", message=FALSE, warning=FALSE}

sdState<-function(fit, y=max(fit$data$years)-1:0){
  idx <- names(fit$sdrep$value) == "logR"
  sdLogR<-fit$sdrep$sd[idx][fit$data$years%in%y]
  idx <- names(fit$sdrep$value) == "logssb"
  sdLogSSB<-fit$sdrep$sd[idx][fit$data$years%in%y]
  idx <- names(fit$sdrep$value) == "logfbar"
  sdLogF<-fit$sdrep$sd[idx][fit$data$years%in%y]
  ret<-cbind(sdLogR, sdLogSSB, sdLogF)
  rownames(ret)<-y
  colnames(ret)<-c("sd(log(R))", "sd(log(SSB))", "sd(log(Fbar))")
  return(ret)
}

pander::pandoc.table(
  as.data.frame(sdState(fit)) %>%  rownames_to_column(var="Year"),
  style="simple", split.tables=200, justify="right", missing=".",
  round=3)

```

\newpage

**Short term forecasts**

Four short term forecasts are presented: 
* Fsq assumption (F2022=F2021=0.185)
* Catch constraint assumption (Catch 2022 = Catch 2021 = 13081 t.)
* TAC constraint assumption (Catch 2022 = 'TAC' 2022 = 29626 t.)
* TAC uptake assumption (Catch 2022 = 'TAC' 2022 * 0.63 = 18873 t.)

In all scenarios, the forecasts starts in 2021 (last data year) with an fsq assumption. The start in the last data year is a requirement of the SAM forecast. 

**Forecast with Fsq assumption (F2022=F2021=0.185)**

```{r, echo=FALSE, fig.asp=0.8, fig.align="center", message=FALSE, warning=FALSE}

fit <- get(paste0("fit_", assessments[2]))

dty  <- max(fit$data$years)        # Last data year
imy  <- dty + 1                    # Intermediate year
fcy  <- imy + 1                    # Forecast year
cty  <- fcy + 1                    # Continuation year
fuy  <- c(imy, fcy, cty)           # Future years

Ry   <- fit$data$years[(length(fit$data$years)-10):(length(fit$data$years)-1)]
fsq  <- stockassessment::fbartable(fit)[nrow(fbartable(fit)),1]
ft   <- c(  fsq, 0.000001,   0.24,    0.33,  seq(0.01,0.3,0.01) )
labs <- c("Fsq",  "Fzero", "Fmsy", "Fp0.5",  paste0("F=",as.character(seq(0.01,0.3,0.01))) )
# ft   <- c( fsq,  0.000001,   0.241,   0.33)
# labs <- c("Fsq",  "Fzero",  "Fmsy", "Fp0.5" )

FC<-list()
fc <- data.frame(stringsAsFactors = FALSE)
for (i in 1:length(ft)) {
  # print(ft[i])
  set.seed(12345)
  
  # forecast
  FC[[i]] <-forecastx(
      fit,
      fscale = c(NA, NA, NA, NA, NA),
      fval = c(fsq, fsq, ft[i], ft[i], ft[i]),
      rec.years = Ry,
      label = labs[i],
      processNoise = FALSE,
      addTSB = TRUE
    )
  
  # as data.frame
  fc <-
    bind_rows(
      fc,
      as.data.frame(attributes(FC[[i]])$tab) %>% 
      rownames_to_column(var="year") %>% 
      mutate(
        ft = ft[i], 
        label = labs[i]
      ) %>% 
      tidyr::pivot_longer(names_to="var1", values_to="data", "fbar:median":"tsb:high") %>% 
      tidyr::separate(var1, into=c("var","type"), sep=":") %>% 
      tidyr::pivot_wider(names_from=type, values_from=data)
    )
}

# last data year
fc %>% 
  filter(year==dty) %>% 
  filter(label=="Fsq") %>% 
  filter(var != "tsb") %>% 
  mutate(ft = ifelse(label=="Fsq",format(as.numeric(ft), digits=3),ft)) %>% 
  mutate(value = paste0(median, " (", low,"-", high,")")) %>% 
  dplyr::select(-median, -low, -high, -ft, -label) %>% 
  pandoc.table(.,
                 style = "simple",
                 split.tables=120, 
                 justify = "right",
                 missing=".")

# current year
fc %>% 
  filter(year==imy) %>% 
  filter(label=="Fsq") %>% 
  filter(var != "tsb") %>% 
  mutate(ft = ifelse(label=="Fsq",format(as.numeric(ft), digits=3),ft)) %>% 
  mutate(value = paste0(median, " (", low,"-", high,")")) %>% 
  dplyr::select(-median, -low, -high) %>% 
  pandoc.table(.,
                 style = "simple",
                 split.tables=120, 
                 justify = "right",
                 missing=".")

# forecast
fc %>% 
  mutate(ft = ifelse(label=="Fzero","0.0",ft)) %>% 
  mutate(ft = ifelse(label=="Fsq",format(as.numeric(ft), digits=3),ft)) %>% 
  
  filter(
    (var %in% c("catch")   & year %in% c(fcy)) |
    (var %in% c("ssb") & year %in% c(fcy, cty))
  ) %>% 
  
  mutate(value = paste0(median, " (", low,"-", high,")")) %>% 
  dplyr::select(-median, -low, -high) %>% 
  
  mutate(var2 = paste(var,year,sep="_")) %>% 
  dplyr::select(-year, -var) %>% 

  tidyr::pivot_wider(names_from=var2, values_from=value) %>% 
  dplyr::select(ft,
              label, 
              paste("catch", fcy, sep="_"),
                                    paste("ssb", fcy, sep="_"),
                                    paste("ssb", cty, sep="_")) %>% 
  pandoc.table(.,
               style = "simple",
               split.tables=120, 
               justify = "right",
               missing=".")
  



```
\newpage

**Catch constraint assumption (Catch 2022 = Catch 2021 = 13081 t.)**

```{r, echo=FALSE, fig.asp=0.8, fig.align="center", message=FALSE, warning=FALSE}

fit <- get(paste0("fit_", assessments[2]))

dty  <- max(fit$data$years)        # Last data year
imy  <- dty + 1                    # Intermediate year
fcy  <- imy + 1                    # Forecast year
cty  <- fcy + 1                    # Continuation year
fuy  <- c(imy, fcy, cty)           # Future years

Ry   <- fit$data$years[(length(fit$data$years)-10):(length(fit$data$years)-1)]
fsq  <- stockassessment::fbartable(fit)[nrow(fbartable(fit)),1]
ft   <- c(  fsq, 0.000001,   0.24,    0.33,  seq(0.01,0.3,0.01) )
labs <- c("CC-Fsq",  "CC-Fzero", "CC-Fmsy", "CC-Fp0.5",  paste0("CC-F=",as.character(seq(0.01,0.3,0.01))) )
# ft   <- c(    fsq,     0.000001,     0.241,       0.33)
# labs <- c("TC-Fsq",  "TC-Fzero", "TC-Fmsy", "TC-Fp0.5")

# Catch constraint 2022
cc <- 13081
# cc <- 20000
  
FC<-list()
fc <- data.frame(stringsAsFactors = FALSE)
for (i in 1:length(ft)) {
  # print(ft[i])
  set.seed(12345)
  
  # forecast
  FC[[i]] <-forecastx(
      fit,
      fscale = c(NA, NA, NA, NA, NA),
      catchval = c(NA , cc,    NA,    NA,    NA),
      fval     = c(fsq, NA, ft[i], ft[i], ft[i]),
      rec.years = Ry,
      label = labs[i],
      processNoise = FALSE,
      addTSB = TRUE
    )

  # as data.frame
  fc <-
    bind_rows(
      fc,
      as.data.frame(attributes(FC[[i]])$tab) %>% 
        rownames_to_column(var="year") %>% 
        mutate(
          ft = get("fbar:median"), 
          label = labs[i]
        ) %>% 
        tidyr::pivot_longer(names_to="var1", values_to="data", "fbar:median":"tsb:high") %>% 
        tidyr::separate(var1, into=c("var","type"), sep=":") %>% 
        tidyr::pivot_wider(names_from=type, values_from=data)
    )
}

      # as.data.frame(attributes(FC[[3]])$tab) %>% 
      # rownames_to_column(var="year")  %>% 
      # mutate(
      #   ft = get("fbar:median"), 
      #   label = labs[i]
      # ) 

# last data year
fc %>% 
  filter(year==dty) %>% 
  filter(label=="CC-Fsq") %>% 
  filter(var != "tsb") %>% 
  mutate(ft = ifelse(label=="Fsq",format(as.numeric(ft), digits=3),ft)) %>% 
  mutate(value = paste0(median, " (", low,"-", high,")")) %>% 
  dplyr::select(-median, -low, -high, -ft, -label) %>% 
  pandoc.table(.,
                 style = "simple",
                 split.tables=120, 
                 justify = "right",
                 missing=".")

        
# current year
fc %>% 
  filter(year==imy) %>% 
  filter(label=="CC-Fsq") %>% 
  filter(var != "tsb") %>% 
  mutate(ft = ifelse(label=="Fsq",format(as.numeric(ft), digits=3),ft)) %>% 
  mutate(value = paste0(median, " (", low,"-", high,")")) %>% 
  dplyr::select(-median, -low, -high, -ft, -label) %>% 
  pandoc.table(.,
                 style = "simple",
                 split.tables=120, 
                 justify = "right",
                 missing=".")

# forecast
fc %>% 
  
  filter(
    (var %in% c("catch") & year %in% c(fcy)) |
    (var %in% c("ssb")   & year %in% c(fcy, cty))
  ) %>% 
  
  mutate(ft = ifelse(label=="Fzero","0.0",ft)) %>% 
  mutate(ft = ifelse(label=="Fsq",format(as.numeric(ft), digits=3),ft)) %>% 
  
  
  mutate(value = paste0(median, " (", low,"-", high,")")) %>% 
  dplyr::select(-median, -low, -high) %>% 
  
  mutate(var2 = paste(var,year,sep="_")) %>% 
  dplyr::select(-year, -var) %>% 

  tidyr::pivot_wider(names_from=var2, values_from=value) %>% 
  dplyr::select(ft,
              label, 
              paste("catch", fcy, sep="_"),
                                    paste("ssb", fcy, sep="_"),
                                    paste("ssb", cty, sep="_")) %>% 
  pandoc.table(.,
               style = "simple",
               split.tables=120, 
               justify = "right",
               missing=".")
  



```

\newpage

**TAC constraint assumption (Catch 2022 = 'TAC' 2022 = 29626 t.)**

```{r, echo=FALSE, fig.asp=0.8, fig.align="center", message=FALSE, warning=FALSE}

fit <- get(paste0("fit_", assessments[2]))

dty  <- max(fit$data$years)        # Last data year
imy  <- dty + 1                    # Intermediate year
fcy  <- imy + 1                    # Forecast year
cty  <- fcy + 1                    # Continuation year
fuy  <- c(imy, fcy, cty)           # Future years

Ry   <- fit$data$years[(length(fit$data$years)-10):(length(fit$data$years)-1)]
fsq  <- stockassessment::fbartable(fit)[nrow(fbartable(fit)),1]
ft   <- c(  fsq, 0.000001,   0.24,    0.33,  seq(0.01,0.3,0.01) )
labs <- c("TC-Fsq",  "TC-Fzero", "TC-Fmsy", "TC-Fp0.5",  paste0("TC-F=",as.character(seq(0.01,0.3,0.01))) )
# ft   <- c(    fsq,     0.000001,     0.241,       0.33)
# labs <- c("TC-Fsq",  "TC-Fzero", "TC-Fmsy", "TC-Fp0.5")

FC<-list()
fc <- data.frame(stringsAsFactors = FALSE)
for (i in 1:length(ft)) {
  # print(ft[i])
  set.seed(12345)
  
  # forecast
  FC[[i]] <-forecastx(
      fit,
      fscale = c(NA, NA, NA, NA, NA),
      catchval = c(NA , 29626,    NA,    NA,    NA),
      fval     = c(fsq,    NA, ft[i], ft[i], ft[i]),
      rec.years = Ry,
      label = labs[i],
      processNoise = FALSE,
      addTSB = TRUE
    )

  # as data.frame
  fc <-
    bind_rows(
      fc,
      as.data.frame(attributes(FC[[i]])$tab) %>% 
        rownames_to_column(var="year") %>% 
        mutate(
          ft = get("fbar:median"), 
          label = labs[i]
        ) %>% 
        tidyr::pivot_longer(names_to="var1", values_to="data", "fbar:median":"tsb:high") %>% 
        tidyr::separate(var1, into=c("var","type"), sep=":") %>% 
        tidyr::pivot_wider(names_from=type, values_from=data)
    )
}


# last data year
fc %>% 
  filter(year==dty) %>% 
  filter(label=="TC-Fsq") %>% 
  filter(var != "tsb") %>% 
  mutate(ft = ifelse(label=="Fsq",format(as.numeric(ft), digits=3),ft)) %>% 
  mutate(value = paste0(median, " (", low,"-", high,")")) %>% 
  dplyr::select(-median, -low, -high, -ft, -label) %>% 
  pandoc.table(.,
                 style = "simple",
                 split.tables=120, 
                 justify = "right",
                 missing=".")

# current year
fc %>% 
  filter(year==imy) %>% 
  filter(label=="TC-Fsq") %>% 
  filter(var != "tsb") %>% 
  mutate(ft = ifelse(label=="Fsq",format(as.numeric(ft), digits=3),ft)) %>% 
  mutate(value = paste0(median, " (", low,"-", high,")")) %>% 
  dplyr::select(-median, -low, -high, -ft, -label) %>% 
  pandoc.table(.,
                 style = "simple",
                 split.tables=120, 
                 justify = "right",
                 missing=".")

# forecast
fc %>% 
  
  filter(
    (var %in% c("catch") & year %in% c(fcy)) |
    (var %in% c("ssb")   & year %in% c(fcy, cty))
  ) %>% 
  
  mutate(ft = ifelse(label=="Fzero","0.0",ft)) %>% 
  mutate(ft = ifelse(label=="Fsq",format(as.numeric(ft), digits=3),ft)) %>% 
  
  
  mutate(value = paste0(median, " (", low,"-", high,")")) %>% 
  dplyr::select(-median, -low, -high) %>% 
  
  mutate(var2 = paste(var,year,sep="_")) %>% 
  dplyr::select(-year, -var) %>% 

  tidyr::pivot_wider(names_from=var2, values_from=value) %>% 
  dplyr::select(ft,
              label, 
              paste("catch", fcy, sep="_"),
                                    paste("ssb", fcy, sep="_"),
                                    paste("ssb", cty, sep="_")) %>% 
  pandoc.table(.,
               style = "simple",
               split.tables=120, 
               justify = "right",
               missing=".")
  



```

\newpage

**TAC uptake assumption (Catch 2022 = 'TAC' 2022 * 0.63 = 18873 t.)**

This forecast is a modified catch constraint based on the regression between TAC and uptake)

```{r, echo=FALSE, fig.asp=0.8, fig.align="center", message=FALSE, warning=FALSE}

fit <- get(paste0("fit_", assessments[2]))

dty  <- max(fit$data$years)        # Last data year
imy  <- dty + 1                    # Intermediate year
fcy  <- imy + 1                    # Forecast year
cty  <- fcy + 1                    # Continuation year
fuy  <- c(imy, fcy, cty)           # Future years

Ry   <- fit$data$years[(length(fit$data$years)-10):(length(fit$data$years)-1)]
fsq  <- stockassessment::fbartable(fit)[nrow(fbartable(fit)),1]
ft   <- c(  fsq, 0.000001,   0.24,    0.33,  seq(0.01,0.3,0.01) )
labs <- c("CC-Fsq",  "CC-Fzero", "CC-Fmsy", "CC-Fp0.5",  paste0("CC-F=",as.character(seq(0.01,0.3,0.01))) )
# ft   <- c(    fsq,     0.000001,     0.241,       0.33)
# labs <- c("TC-Fsq",  "TC-Fzero", "TC-Fmsy", "TC-Fp0.5")

# Catch constraint 2022
cc <- 18873
# cc <- 20000

FC<-list()
fc <- data.frame(stringsAsFactors = FALSE)
for (i in 1:length(ft)) {
  # print(ft[i])
  set.seed(12345)
  
  # forecast
  FC[[i]] <-forecastx(
      fit,
      fscale = c(NA, NA, NA, NA, NA),
      catchval = c(NA , cc,    NA,    NA,    NA),
      fval     = c(fsq, NA, ft[i], ft[i], ft[i]),
      rec.years = Ry,
      label = labs[i],
      processNoise = FALSE,
      addTSB = TRUE
    )

  # as data.frame
  fc <-
    bind_rows(
      fc,
      as.data.frame(attributes(FC[[i]])$tab) %>% 
        rownames_to_column(var="year") %>% 
        mutate(
          ft = get("fbar:median"), 
          label = labs[i]
        ) %>% 
        tidyr::pivot_longer(names_to="var1", values_to="data", "fbar:median":"tsb:high") %>% 
        tidyr::separate(var1, into=c("var","type"), sep=":") %>% 
        tidyr::pivot_wider(names_from=type, values_from=data)
    )
}

      # as.data.frame(attributes(FC[[3]])$tab) %>% 
      # rownames_to_column(var="year")  %>% 
      # mutate(
      #   ft = get("fbar:median"), 
      #   label = labs[i]
      # ) 

        
# last data year
fc %>% 
  filter(year==dty) %>% 
  filter(label=="CC-Fsq") %>% 
  filter(var != "tsb") %>% 
  mutate(ft = ifelse(label=="Fsq",format(as.numeric(ft), digits=3),ft)) %>% 
  mutate(value = paste0(median, " (", low,"-", high,")")) %>% 
  dplyr::select(-median, -low, -high, -ft, -label) %>% 
  pandoc.table(.,
                 style = "simple",
                 split.tables=120, 
                 justify = "right",
                 missing=".")

# current year
fc %>% 
  filter(year==imy) %>% 
  filter(label=="CC-Fsq") %>% 
  filter(var != "tsb") %>% 
  mutate(ft = ifelse(label=="Fsq",format(as.numeric(ft), digits=3),ft)) %>% 
  mutate(value = paste0(median, " (", low,"-", high,")")) %>% 
  dplyr::select(-median, -low, -high, -ft, -label) %>% 
  pandoc.table(.,
                 style = "simple",
                 split.tables=120, 
                 justify = "right",
                 missing=".")

# forecast
fc %>% 
  
  filter(
    (var %in% c("catch") & year %in% c(fcy)) |
    (var %in% c("ssb")   & year %in% c(fcy, cty))
  ) %>% 
  
  mutate(ft = ifelse(label=="Fzero","0.0",ft)) %>% 
  mutate(ft = ifelse(label=="Fsq",format(as.numeric(ft), digits=3),ft)) %>% 
  
  
  mutate(value = paste0(median, " (", low,"-", high,")")) %>% 
  dplyr::select(-median, -low, -high) %>% 
  
  mutate(var2 = paste(var,year,sep="_")) %>% 
  dplyr::select(-year, -var) %>% 

  tidyr::pivot_wider(names_from=var2, values_from=value) %>% 
  dplyr::select(ft,
              label, 
              paste("catch", fcy, sep="_"),
                                    paste("ssb", fcy, sep="_"),
                                    paste("ssb", cty, sep="_")) %>% 
  pandoc.table(.,
               style = "simple",
               split.tables=120, 
               justify = "right",
               missing=".")
  


# x <-hcr(
#     fit,
#     nYears = 3,
#     Ftarget = 0.24,
#     Btrigger = 82999,
#     Blim = 0,
#     preForecast = list(catchval = 30000),
#     rec.years = Ry
#   )
#   
# as.data.frame(attributes(x$forecast)$tab) %>% 
#   rownames_to_column(var="year") %>% 
#   mutate(
#     ft = get("fbar:median"), 
#     label = labs[i]
#   ) %>% 
#   tidyr::pivot_longer(names_to="var1", values_to="data", "fbar:median":"catch:high") %>% 
#   tidyr::separate(var1, into=c("var","type"), sep=":") %>% 
#   tidyr::pivot_wider(names_from=type, values_from=data) %>% 
#   filter(
#     (var %in% c("catch") & year %in% c(fcy)) |
#     (var %in% c("ssb")   & year %in% c(fcy, cty))
#   ) %>% 
#   
#   mutate(ft = ifelse(label=="Fzero","0.0",ft)) %>% 
#   mutate(ft = ifelse(label=="Fsq",format(as.numeric(ft), digits=3),ft)) %>% 
#   
#   
#   mutate(value = paste0(median, " (", low,"-", high,")")) %>% 
#   dplyr::select(-median, -low, -high) %>% 
#   
#   mutate(var2 = paste(var,year,sep="_")) %>% 
#   dplyr::select(-year, -var) %>% 
# 
#   tidyr::pivot_wider(names_from=var2, values_from=value) %>% 
#   dplyr::select(ft,
#               label, 
#               paste("catch", fcy, sep="_"),
#                                     paste("ssb", fcy, sep="_"),
#                                     paste("ssb", cty, sep="_")) %>% 
#   pandoc.table(.,
#                style = "simple",
#                split.tables=120, 
#                justify = "right",
#                missing=".")

  
```



\newpage

**FIGURES**

**Catch numbers (Crayola)**

```{r, echo=FALSE, fig.asp=1.3, fig.align="center", message=FALSE, warning=FALSE}

t <-
  opr %>% 
  filter(file==nam) %>%
  filter(!is.na(age)) %>% 

  complete(file, fleet, year, age) %>% 
  group_by(file, fleet, year, age) %>% 
  summarise(oC = sum(exp(o), na.rm=TRUE)) %>% 
  group_by(file, fleet, age) %>% 
  mutate(value = oC/mean(oC, na.rm=TRUE)) %>% 
  mutate(yc = year - age) %>% 
  data.frame()
  
for (i in 1:length(unique(t$fleet))) {
  assign(paste0("p", i),
         t %>% 
           filter(fleet == sort(unique(t$fleet))[i]) %>% 
          
           ggplot() +
           theme_bw() +
           theme(legend.position = "none") +
           theme(axis.text.y     = element_blank()) +
           theme(panel.border    = element_rect(colour="black" , size=0.1)) +
           theme(axis.ticks.y    = element_blank() ) +
           theme(axis.text.x     = element_text(angle = 90, vjust = 0.5, hjust=1, size=10)) +
           theme(panel.spacing   = unit(0.2, "lines")) +
           theme(plot.margin=unit(c(0,0,0,0),"mm")) +
           geom_col(aes(year, value, fill = factor(yc))) + 
           ggmisc::scale_fill_crayola() +
           scale_x_continuous(breaks=seq(min(t$year), max(t$year), 1)) +
           labs(x = NULL, y = NULL, title=NULL) +
           facet_grid(age ~ fleet, scale = "free_y", switch = "y") 
      )
  
}

print(cowplot::plot_grid(p1, p2, p3,
                   nrow=1, scale=0.98, align="hv",
                   rel_heights = c(1.5, 1.5, 1.5, 1.5)))



```


\newpage

**Catch weights**

```{r, echo=FALSE, fig.asp=0.8, fig.align="center", message=FALSE, warning=FALSE}

t <-
  rbya %>% 
  filter(file==nam) %>%
  filter(cW != -1) %>% 
  mutate(cW = as.integer(cW * 1000)) 

tt <-
  t %>% 
  group_by(file, age) %>% 
  slice_tail(n=1)

t %>%   
  ggplot(aes(x=year, y=cW, group=age)) +
  theme_publication() +
  theme(legend.position = "none") +
  geom_point(aes(colour=as.character(age))) +
  geom_line(aes(colour=as.character(age))) +
  ggrepel::geom_text_repel(data=tt, aes(x=year, y=cW, label=as.character(age), colour=as.character(age)), 
                           hjust=0, direction="y", nudge_x = 1, segment_size=0.4, force=0.2,
                           segment.colour="gray", box.padding = 0.5, min.segment.length = 0 ) +
  scale_x_continuous(expand = expansion(mult = c(0.02 , 0.1))) +
  labs(x="", y="cW (gr)") +
  expand_limits(y=0) 

t %>%   
  ggplot(aes(x=year, y=cW, group=age)) +
  theme_publication() +
  theme(axis.text.x     = element_text(angle = 90, vjust = 0.5, hjust=1, size=10)) +
  theme(legend.position = "none") +
  geom_point(aes(colour=as.character(age))) +
  geom_line(aes(colour=as.character(age))) +
  expand_limits(y=0) +
  scale_x_continuous(breaks=scales::pretty_breaks()) +
  labs(x="", y="cW (gr)") +
  facet_wrap(~age, scales="free_y")

```

\newpage

**Comparison of the assessment with two different settings**

Comparison of the assessment with two different settings: 

1. par$logSdLogN<-c(0,log(0.0001)) [assessment 2021]
2. par$logSdLogN<-c(0,log(0.01)) [assessment 2022]

No noticeable difference between the two settings. Note that the retrospective analysis in 2022 could not be run with settings as in WG2021 but did work with the updated settings. 

```{r, echo=FALSE, fig.asp=0.9, fig.align="center", message=FALSE, warning=FALSE}

# # data
# fil <- file.path(filepath, "data.RData")
# if(RCurl::url.exists(fil)) {
#   load(url(fil))
# }
# 
# # config
# fil <- file.path(configpath, "model.cfg")
# if(RCurl::url.exists(fil)) {
#   conf <- loadConf(dat, url(fil), patch=TRUE)
# }
# 
# par<-defpar(dat,conf)
# 
# par$logSdLogN<-c(0,log(0.01))
# fit1<-sam.fit(dat,conf,par, map=list(logSdLogN=factor(c(1,NA))))
# if(fit1$opt$convergence!=0) stop("Model did not converge.")
# rby1 <- fishvice::sam_rby(fit1) %>% mutate(model="1")
# 
# par$logSdLogN<-c(0,log(0.0001))
# fit2<-sam.fit(dat,conf,par, map=list(logSdLogN=factor(c(1,NA))))
# if(fit2$opt$convergence!=0) stop("Model did not converge.")
# rby2 <- fishvice::sam_rby(fit2) %>% mutate(model="2")

# save(dat, conf, fit1, rby1, fit2, rby2, file=file.path(localpath, "rerun.RData"))
load(file=file.path(localpath, "rerun.RData"))

bind_rows(rby1, rby2) %>% 
  # filter(variable == "ssb") %>% 
  ggplot(aes(x=year, y=est, colour=model)) +
  geom_line() +
  geom_ribbon(aes(xmin = year, xmax=year, ymin=low, ymax=high, fill=model), alpha=0.4) +
  facet_wrap(~variable, scales="free_y")


```

\newpage

**Observed and predicted (logscale): catch**

```{r, echo=FALSE, fig.asp=1.3, fig.align="center", message=FALSE, warning=FALSE}

# generate plots for observations and predictions

t <-
  opr %>%
  filter(file==nam) %>% 
  rename(obs=o, pred=p) %>% 
  mutate(
    fleet = ifelse(grepl("summer"       , tolower(fleet)), "FSS",  fleet),
    fleet = ifelse(grepl("faroesedeep"  , tolower(fleet)), "FDWS", fleet),
    fleet = ifelse(grepl("scottishdeep" , tolower(fleet)), "SDWS", fleet),
    fleet = ifelse(grepl("cpue"         , tolower(fleet)), "CPUE", fleet)
  ) %>% 
  mutate(
    label = paste(fleet, ifelse(is.na(age), "biomass", "age"), ifelse(is.na(age), "", stringr::str_pad(age, width=2, pad="0")))
  )

# t1 <- t %>% filter(!is.na(age)) %>% distinct(fleet)
# t2 <- t %>% filter(is.na(age)) %>% distinct(fleet)

t %>% 
  ggplot(aes(x=year, y=obs)) +
  theme_publication() +
  geom_point(colour="red", shape=21) +
  geom_line(aes(y=pred), colour="black") +
  scale_x_continuous(breaks=scales::pretty_breaks(n=3)) +
  labs(title="", x="", y="log") +
  facet_wrap(~label, ncol=5, scales="free_y")


```

FSS = Faroese Summer Survey

FDWS = Faroese Deepwater Survey

SDWS = Scottish Deepwater Survey

CPUE = Faroese EU CPUE index

\newpage


**One step ahead residuals**

```{r echo=FALSE, fig.align="center", fig.asp=1.2, message=FALSE, warning=FALSE}

plot(get(paste0("res_", assessments[2])))

```
\newpage

**Process residuals**

```{r echo=FALSE, fig.align="center", fig.asp=0.8, message=FALSE, warning=FALSE}

plot(get(paste0("procres_", assessments[2])))

```

\newpage

**Parameter estimates**

```{r echo=FALSE, fig.align="center", fig.asp=1.2, message=FALSE, warning=FALSE}

# **Parameter estimates**

plist <- list()

getparsfromfit <- function(fit) {
  # a  <- gsub("fit_", "", fit) 
  df <- 
    data.frame(stockassessment::partable(fit)) %>% 
    add_rownames() %>%  
    separate(rowname, into=c("parameter", "parnumber"), sep="_")
  return(df)
}

pars2 <- tibble()
for (a in assessments) {
  assign(paste0("pars2_",a), mutate(getparsfromfit(get(paste0("fit_", a))), file=a))  
  pars2 <- 
    bind_rows(
      pars2, 
      get(paste0("pars2_",a)) %>% mutate(parnumber = factor(parnumber, levels=unique(.$parnumber))) %>% lowcase())
}
 

# plot
plist[[1]] <-
  pars2 %>% 
  filter(tolower(parameter) %in% c("logfpar", "logsdlogfsta","logsdlogn", "logsdlogobs","transfirardist")) %>%

  ggplot(aes(x=parnumber, y=exppar, group=file)) +
  
  theme_publication() +
  theme(legend.title=element_blank()) +
  theme(legend.position="none") +
  # theme(axis.text.x     = element_text(angle = 90, vjust = 0.5, hjust=1, size=10)) +
  theme(strip.text.y.right = element_text(angle = 0)) +
  theme(plot.margin=grid::unit(c(0,0,0,0), "mm")) +
  
  geom_bar(aes(fill=file), stat="identity", position="dodge") +
  geom_errorbar(aes(ymin=low, ymax=high, colour=file, group=file), position=position_dodge()) +
  geom_point(aes(colour=file), position=position_dodge(0.9)) +
  scale_colour_manual(values=c( "#56B4E9", "#999999", "#E69F00"))+
  coord_flip()+
  labs(x="",y="") +
  facet_grid(parameter~., scales="free",space = "free")

plist[[2]] <-
  pars2 %>%
  filter(tolower(parameter) %in% c("itrans")) %>%

  ggplot(aes(x=parnumber, y=exppar, group=file)) +
  
  theme_publication() +
  theme(legend.title=element_blank()) +
  theme(legend.position="none") +
  # theme(axis.text.x     = element_text(angle = 90, vjust = 0.5, hjust=1, size=10)) +
  theme(strip.text.y.right = element_text(angle = 0)) +
  theme(plot.margin=grid::unit(c(0,0,0,0), "mm")) +
  
  geom_bar(aes(fill=file), stat="identity", position="dodge") +
  geom_errorbar(aes(ymin=low, ymax=high, colour=file, group=file), position=position_dodge()) +
  geom_point(aes(colour=file), position=position_dodge(0.9)) +
  scale_colour_manual(values=c( "#56B4E9", "#999999", "#E69F00"))+
  coord_flip()+
  labs(x="",y="") +
  facet_grid(parameter~., scales="free",space = "free")

print(patchwork::wrap_plots(plist) +
  patchwork::plot_layout(ncol = 1, heights = c(0.92, 0.08)) +
  patchwork::plot_annotation() & theme(plot.tag = element_text(size = 8)))

```


\newpage

**Comparisons of the assessment with previous year assessment**

```{r, echo=FALSE, fig.asp=0.8, fig.align="center", message=FALSE, warning=FALSE}

# plot of trend by variable
rby %>% 
  # filter(stock %in% c("ARU.27.5b6a_WGDEEP_2020_","ARU.27.5b6a_WGDEEP_2020_mp")) %>% 
  filter(variable %in% c("catch", "ssb","fbar","rec")) %>% 
  mutate(variable = factor(variable, levels=c("catch", "ssb","fbar","rec"))) %>% 
  mutate(year = ifelse(year < 1000, year+1994, year)) %>% 
  
  ggplot(aes(x=year,y=est, group=file)) +
  theme_publication() +
  theme(legend.title=element_blank()) +
  geom_ribbon(aes(ymin=low, ymax=high, fill=file), alpha=0.3) +
  geom_line(aes(colour=file)) +
  geom_point(aes(colour=file)) +
  
  geom_point(data=obscatch, aes(x=year,y=est), inherit.aes = FALSE, shape=3) +
  
  scale_y_continuous(labels = addUnits) +
  expand_limits(y=0) +
  labs(x="", y="") +
  facet_wrap(~variable, scales="free_y", ncol=2)

# stockassessment::fbartable(fit)


```

\newpage


**Summary of the assessment (figure)**

```{r, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}


# plot of trend by variable
rby %>% 
  filter(file %in% assessments[2]) %>% 
  filter(variable %in% c("ssb","fbar","rec")) %>% 
  mutate(variable = factor(variable, levels=c("ssb","fbar","rec"))) %>% 
  
  ggplot(aes(x=year,y=est, group=file)) +
  theme_publication() +
  theme(legend.title=element_blank()) +
  geom_ribbon(aes(ymin=low, ymax=high, fill=file), alpha=0.3) +
  geom_line(aes(colour=file)) +
  geom_point(aes(colour=file)) +
  geom_hline(data=filter(rp, grepl("msy", refpoint)), aes(yintercept=value), linetype="dashed") +
  geom_hline(data=filter(rp, grepl("lim", refpoint)), aes(yintercept=value)) +
  
  geom_hline(data=filter(rp, grepl("msy", refpoint)), aes(yintercept=value), linetype="dashed") +
  geom_hline(data=filter(rp, grepl("lim", refpoint)), aes(yintercept=value)) +
  
  scale_y_continuous(labels = addUnits) +
  expand_limits(y=0) +
  labs(x="", y="") +
  facet_wrap(~variable, scales="free_y", ncol=1)

# stockassessment::fbartable(fit)


```

\newpage

**Retro of the Assessments**

```{r, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}

rho <- 
  mohn(get(paste("retro", assessments[2], sep="_"))) %>% 
  as.data.frame() %>% 
  setNames("rho") %>% 
  rownames_to_column(var="variable") %>% 
  mutate(
    variable = case_when(
      grepl("^R", variable) ~ "rec",
      grepl("^S", variable) ~ "ssb",
      grepl("^F", variable) ~ "fbar"
    )
  ) %>% 
  mutate(rho = round(rho, digits=3)) %>% 
  mutate(variable = factor(variable, levels=c("ssb","fbar","rec")))


t <-
  retro_rby %>% 
  filter(variable %in% c("ssb","fbar","rec")) %>% 
  filter(file == assessments[[2]]) %>% 
  ungroup() %>% 
  mutate(plottype = ifelse(finalyear == max(finalyear), "last", NA)) %>% 
  mutate(finalyear = as.character(finalyear)) %>% 
  mutate(variable = factor(variable, levels=c("ssb","fbar","rec")))
  
# plot of trend by variable
t %>% 
  ggplot(aes(x=year,y=est, group=finalyear)) +
  theme_publication() +
  theme(legend.position = "none") +
  theme(legend.title=element_blank()) +
  
  # last year
  geom_ribbon(data=filter(t, !is.na(plottype)), 
              aes(ymin=low, ymax=high), fill="gray", alpha=0.3) +
  geom_line(data=filter(t, !is.na(plottype)),
            colour="black", size=1) +
  
  # othyer years
  geom_line(data=filter(t, is.na(plottype)),
            aes(colour=finalyear),
            size=0.8) +

  # geom_point(aes(colour=finalyear)) +
  scale_y_continuous(labels = addUnits) +
  
  geom_text(data=rho, aes(label=paste("mohn=", rho), x=-Inf, y=Inf), hjust=0, vjust=1, inherit.aes = FALSE) +
  expand_limits(y=0) +
  labs(x="", y="") +
  facet_wrap(~variable, scales="free_y", ncol=1)

# stockassessment::fbartable(fit)


```

\newpage

**Leave one out runs**

```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

t <-
  loo_rby %>% 
  filter(variable %in% c("ssb","fbar","rec")) %>% 
  filter(file == assessments[[2]]) %>% 
  mutate(variable = factor(variable, levels=c("ssb","fbar","rec")))
  
x <-
  get(paste("rby", assessments[2], sep="_")) %>% 
  filter(variable %in% c("ssb","fbar","rec")) %>% 
   mutate(variable = factor(variable, levels=c("ssb","fbar","rec")))


# plot of LOO
t %>% 
  ggplot(aes(x=year,y=est, group=leftout)) +
  theme_publication() +
  # theme(legend.position = "none") +
  theme(legend.title=element_blank()) +
  
  geom_ribbon(data=x,aes(x=year, ymin=low, ymax=high), fill="gray", alpha=0.3, inherit.aes = FALSE) +
  geom_line(data=x,aes(x=year, y=est), colour="black", size=1, inherit.aes = FALSE) +
  
  geom_line(aes(colour=leftout), size=0.8) +
  # geom_point(aes(colour=leftout)) +
  scale_y_continuous(labels = addUnits) +
  expand_limits(y=0) +
  labs(x="", y="") +
  facet_wrap(~variable, scales="free_y", ncol=1)

# stockassessment::fbartable(fit)


```






```{r eval=FALSE, fig.align="center", fig.asp=0.9, message=FALSE, warning=FALSE, include=FALSE}

# NOT USED 

# generate plots for observations and predictions

t <-
  opr %>%
  filter(file==nam) %>% 
  rename(obs=o, pred=p) 

t1 <- t %>% filter(!is.na(age)) %>% distinct(fleet)
t2 <- t %>% filter(is.na(age)) %>% distinct(fleet)

# age based residuals
# f <- t1$fleet[1]
for (f in t1$fleet) {
  p <-
    t %>% 
    filter(fleet == f) %>% 
    
    ggplot(aes(x=year, y=obs)) +
    theme_publication() +
    geom_point(colour="red", shape=21) +
    geom_line(aes(y=pred), colour="black") +
    # scale_y_continuous(limits = c(-max(abs(t$res), na.rm=TRUE), max(abs(t$res), na.rm=TRUE))) +
    scale_x_continuous(breaks=scales::pretty_breaks(n=3)) +
    labs(title=f, x="", y="log") +
    facet_wrap(~age, ncol=5)
  
  print(p)
  
}

# biomass residuals
t %>% 
  filter(fleet %in% t2$fleet) %>% 
  mutate(age = "biomass") %>% 

  ggplot(aes(x=year, y=obs)) +
  theme_publication() +
  geom_point(colour="red", shape=21) +
  geom_line(aes(y=pred), colour="black") +
  scale_x_continuous(breaks=scales::pretty_breaks(n=3)) +
  labs(x="", y="log") +
  facet_wrap(~fleet, nrow=1, scales="free_y")
  


```


```{r eval=FALSE, fig.align="center", fig.asp=1.2, message=FALSE, warning=FALSE, include=FALSE}

# NOT USED

# **Parameter estimates by fleet**

pars %>% 
  mutate(fleet = ifelse(grepl("Combine", fleet), "Faroe-EU CPUE", fleet)) %>% 
  ggplot(aes(x=age, y=est)) +
  
  theme_publication() +
  theme(legend.title=element_blank()) +
  theme(legend.position="none") +
  # theme(axis.text.x     = element_text(angle = 90, vjust = 0.5, hjust=1, size=10)) +
  theme(strip.text.y.right = element_text(angle = 0)) +
  theme(plot.margin=grid::unit(c(0,0,0,0), "mm")) +
  
  geom_bar(aes(fill=file), stat="identity", position="dodge") +
  geom_errorbar(aes(ymin=low, ymax=high, colour=file, group=file), position=position_dodge()) +
  geom_point(aes(colour=file), position=position_dodge(0.9)) +
  scale_colour_manual(values=c( "#56B4E9", "#999999", "#E69F00"))+
  # coord_flip()+
  labs(x="",y="") +
  facet_grid(what~fleet, scales="free_y")

```