---
output: 
  word_document:
    reference_docx: ../report_template_v1.6.dotx
---

```{r setup, include=FALSE}
# -------------------------------------------------------------------------------------
# CPUE standardizing for argentines: pfa + faroer data
#
# 23/04/2020 modified the calculation of CPUE (catch by day and by rectangle)
# 25/04/2020 added option for standardization by haul; rearranged data processing steps
# 09/04/2021 adapted for WGDEEP 2021
# -------------------------------------------------------------------------------------

# ================================================================================
# 0. Initialization. Set working directory and filename and sourcing
# ================================================================================

require("knitr")
knitr::opts_chunk$set(echo=FALSE,fig.width=10,message=FALSE,warning=FALSE,comment="",crop=TRUE)
knit_hooks$set(crop = hook_pdfcrop)

rm(list=ls())
options(dplyr.summarise.inform = FALSE)

global_plot_map   <- TRUE
global_plot_fao   <- TRUE
global_plot_depth <- FALSE

# Libraries
library(rmarkdown)
library(maps)          # world maps
library(sp)            # spatial data 
require(geosphere, quietly=TRUE)           # spatial manipulation

library(pander)        # tables
library(colorspace)    # hcl colours
library(viridis)       # Viridis colour scheme
library(RColorBrewer)  # ColourBrewer

library(tidyverse)     # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
library(lubridate)     # data handling
library(reshape2)      # reshaping data; e.g. cast
library(readxl)        # excel reader
library(broom)         # clean up statistics
library(scales)        # pretty scales
library(stringr)       # string manipulations
library(magrittr)      # for e.g. set_colnames
library(captioner)     # captioning of figures and tables

library(mgcv)          # tensor spline for gam
library(lme4)
library(MASS)
library(mgcViz)         # GAM output to ggplot

# To number figures and tables
fig_nums <- captioner::captioner(prefix = "Figure ")
tab_nums <- captioner::captioner(prefix = "Table ")

# default settings for tables
panderOptions('table.alignment.default', function(df) ifelse(sapply(df, is.numeric), 'right', 'left'))

# Source helper function
source("lowcase.r")
source("theme_publication.r")
source("get_onedrive.r")

# set onedrive directory
onedrive <- get_onedrive() 

# lastyear, thisyear
lastyear <- 2020
thisyear <- 2021

# set data path
lastyearpath <- "//community.ices.dk@SSL/DavWWWRoot/ExpertGroups/WGDEEP/2020 Meeting Docs/06. Data/aru.27.5b6a/CPUE"
thisyearpath <- "//community.ices.dk@SSL/DavWWWRoot/ExpertGroups/WGDEEP/2021 Meeting Docs/06. Data/aru.27.5b6a/CPUE"

pfa_file    <- "pfa_arg_raw.RData"
faroe_file  <- "FaroeGSS_CommercialTrawlData1995to2020.xlsx"

# wkgsspath  <- "//community.ices.dk/ExpertGroups/benchmarks/2020/wkdeep/2014 Meeting docs/06. Data/aru.27.5b6a/CPUE"
  
# load spatial datasets
load(file.path(onedrive, "rdata/world.df.RData"))
load(file.path(onedrive, "rdata/fao.df.RData"))
load(file.path(onedrive, "rdata/depth200.df.RData"))
load(file.path(onedrive, "rdata/fao.RData"))

rect <- 
  get(load(file.path(onedrive, "rdata/icesrectangles.df.RData"))) %>% 
  group_by(rect) %>% 
  filter(row_number() == 1) %>% 
  dplyr::select(rect, long, lat)


# CI function
make_ci <- function(pred, data){
  fit <- pred$fit
  lwr <- exp(pred$fit - (1.96 * pred$se.fit))
  upr <- exp(pred$fit + (1.96 * pred$se.fit))
  return(data.frame(fit, lwr, upr, data))
  }

# WKGSS 2020 CPUE index
lastyear_cpue_index <-
  read.csv(file.path(lastyearpath, "Standardized commercial cpue 2005-2019.csv")) %>% 
  mutate(type=paste0("wgdeep",lastyear))

# --------------------------------------------------------------------
# read in information as used in this meeting of WGDEEP
# --------------------------------------------------------------------

#  Load the PFA CPUE data
pfa_arg_raw <- 
  get(load(file.path(thisyearpath, pfa_file))) %>% mutate(fleet="pfa") 
# summarize(pfa_arg_raw, nhauls=n())

# Load the Faroer CPUE data

faroer_aru_raw <-
    read_excel(file.path(thisyearpath, faroe_file), 
             col_names = TRUE,
             col_types = "text") %>% 
  lowcase() %>% 
  setNames(gsub("icesrect","rect", names(.))) %>% 
  setNames(gsub("hours","duration", names(.))) %>% 
  setNames(gsub("gsskg","catch", names(.))) %>% 
  setNames(gsub("depthm","depth", names(.))) %>% 
  
  mutate_at(c("vessel", "year", "month", "day"), list(as.integer)) %>% 
  mutate_at(c("duration", "catch", "depth"), list(as.numeric)) %>% 

  mutate(
    catch = catch/1000,   # catch in tonnes
    date  = as.Date(day, origin="1899-12-30"),
    day   = day(date),
    yday  = yday(date),
    week  = as.numeric(week(date)),
    time  = format(strptime(sprintf("%04d", as.numeric(shoottime)), format="%H%M"), 
                   format = "%H:%M"),
    shoot_time = ymd_hm(paste(date, time)), 
    fleet = "faroer", species="aru") %>% 
  
  ungroup() %>% 
  dplyr::select(vesselid=vessel, date, year, month, week, day, yday, shoot_time, depth,
                catch, duration, rect, species, fleet)

comb_arg_byhaul <-
  bind_rows(faroer_aru_raw, pfa_arg_raw) %>% 
  
  # add rectangle information
  left_join(rect, by = "rect") %>% 
  
  # add hset time and depth category 
  mutate(hset      = as.numeric(strftime(shoot_time, format="%H"))) %>% 
  mutate(hset_cat  = cut(hset, include.lowest = TRUE, breaks=c(0,2,4,6,8,10,12,14,16,18,20,22,24))) %>% 
  mutate(depth_cat = cut(depth, breaks=c(0, 200, 300, 400, 500, 600, 700, 2500), right = FALSE)) %>% 
  
  mutate(cpue_hour = catch / duration) %>% 
  mutate(cpue_haul = catch) %>% 
  
  # filter area and weeks
  filter(long > -18, long < 5) %>% 
  # filter(lat  >=59.5, lat <= 62.5) %>% 
  filter(week %in% c(15:34)) %>% 
  filter(year > 2004) %>% 
  filter(!is.na(depth)) 

# comb_arg_byhaul %>% filter(fleet=="pfa") %>% View()
# pfa_arg_raw %>% left_join(rect) %>% dplyr::select(lat) %>% hist(as.numeric(lat))
# comb_arg_byhaul %>%  filter(fleet=="pfa") %>% ggplot(aes(lat)) + geom_histogram()

comb_arg_byday <-
  comb_arg_byhaul %>% 
  
  # aggregate within vessels and days
  group_by(fleet, vesselid, year, week, rect, long, lat) %>%
  summarize(catch    = sum(catch, na.rm=TRUE), 
            duration = sum(duration, na.rm=TRUE),
            nhauls   = n(),
            ndays    = n_distinct(date),
            depth    = mean(depth, na.rm=TRUE)) %>% 
  arrange(fleet, year, week, vesselid) %>% 
  # filter(fleet=="pfa")

  # sum over vessels and days
  group_by(fleet, year, week, rect, long, lat) %>%
  summarize(catch    = sum(catch, na.rm=TRUE), 
            duration = sum(duration, na.rm=TRUE),
            nhauls   = sum(nhauls, na.rm=TRUE),
            ndays    = sum(ndays, na.rm=TRUE),
            depth    = mean(depth, na.rm=TRUE)) %>% 
  
  # calculate CPUE
  mutate(cpue_day = catch / ndays) %>% 
  mutate(depth_cat= cut(depth, breaks=c(0, 200, 300, 400, 500, 600, 700, 2500), right = FALSE)) %>% 
  
  ungroup() %>% 
  
  # explanatory variables should be factors
  mutate(year = as.factor(year),
         week = as.factor(week),
         long_cat = as.factor(long),
         lat_cat = as.factor(lat)) 

# comb_arg_byday %>% filter(fleet=="pfa") %>% View()

```

WGDEEP 2021, WD xx


**CPUE Standardization of Silver smelt in 5b and 6a**

M.A. Pastoors, L.H. Ofstad and H. Olsen

Corresponding author: mpastoors@pelagicfish.eu

`r format(Sys.time(), '%d/%m/%Y')`

**Abstract ** 

At the WKGSS 2020 benchmark of Greater silver smelt in 5b and 6a, a combined and standardized CPUE series for the Faroe and EU fleets has been introduced (Quirijns and Pastoors 2020). On checking the data in preparation for WGDEEP 2020, a small error was detected in the way CPUE was calculated. This report provides a summary of the issue and proposed a method to repair the situation. The overall trend in CPUE is still similar although there are some differences in the most recent year. 


<!--1. Introduction ------------------------------------------------------ -->

# Introduction

At the WKGSS 2020 benchmark of Greater silver smelt in 5b and 6a, a combined and standardized CPUE series for the Faroe and EU fleets has been introduced (Quirijns and Pastoors 2020). During WGDEEP 2020 two small errors were detected in the way CPUE was calculated and solutions to these errors were provided (Pastoors and Quirijns 2020).  This report provides a an update of the CPUE calculation for Greater silver smelt, with the time series update to 2020 according to the method agreed in WGDEEP 2020. 

# Results

```{r echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}

fig_nums(
  name    = "metrics", 
  caption = "ARU.27.5b6a metrics describing the fisheries",
  display = FALSE)

t1 <- comb_arg_byhaul %>% 
  group_by(fleet, year) %>% 
  summarise(nvessels = n_distinct(vesselid)) %>% 
  pivot_longer(names_to="variable", values_to="value", c("nvessels"))

t2 <- comb_arg_byhaul %>% 
  group_by(fleet, year,vesselid) %>% 
  summarise(ndays = n_distinct(date)) %>% 
  group_by(fleet, year) %>% 
  summarise(ndays=sum(ndays, na.rm=TRUE)) %>% 
  pivot_longer(names_to="variable", values_to="value", c("ndays"))

t3 <- comb_arg_byhaul  %>%
  group_by(fleet, year) %>% 
  summarise(
    nhauls = n(),
    catch  = sum(catch, na.rm=TRUE),
    nhours = sum(duration, na.rm=TRUE)) %>% 
  pivot_longer(names_to="variable", values_to="value", c("nhauls","catch","nhours"))

t4 <- comb_arg_byday  %>%
  mutate(year = as.numeric(as.character(year))) %>% 
  group_by(fleet, year) %>% 
  summarise(
    catch_day = sum(catch, na.rm=TRUE)/sum(ndays, na.rm=TRUE)) %>% 
  pivot_longer(names_to="variable", values_to="value", c("catch_day"))

bind_rows(t1, t2, t3, t4) %>% 
  mutate(variable = factor(variable, levels=c("nvessels","ndays","nhauls","nhours","catch","catch_day"))) %>% 
  ggplot(aes(x=year, y=value, group=fleet)) +
  theme_publication() +
  geom_bar(aes(fill=fleet), stat="identity", position = position_dodge2(preserve = "single") ) +
  labs(x="", y="") +
  facet_wrap(~variable, scales="free_y", ncol=2)



```
*`r fig_nums("metrics")`*

##### page break

The 'raw' (unstandardized) CPUE is based on the catch per day. 

```{r echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

fig_nums(
  name    = "cpue", 
  caption = "ARU.27.5b6a Catch per unit effort.",
  display = FALSE)

comb_arg_byday %>%
  mutate(
    year      = factor(year),
    catch_day = cpue_day) %>% 
    
  ggplot(aes(x=year, y=catch_day)) +
  theme_publication() +
  theme(legend.position="none") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  geom_boxplot(aes(fill=factor(fleet)), position = position_dodge2(preserve = "single")) +
  labs(x="") +
  facet_wrap(~fleet, scales="free_y")


```

*`r fig_nums("cpue")`*

##### page break

For the years 2005-`r lastyear`, below are the spatial distributions of the used number of hauls by fleet. 

```{r echo=FALSE, fig.asp=0.8, fig.align="center", message=FALSE, warning=FALSE}

fig_nums(
  name    = "plothauls", 
  caption = "ARU.27.5b6a plot of the number of hauls by rectangle and day",
  display = FALSE)

bind_rows(pfa_arg_raw, faroer_aru_raw) %>% 
  left_join(rect) %>% 
  mutate(year = as.numeric(as.character(year))) %>% 
  group_by(fleet, year, lat, long) %>% 
  summarize(nhauls = n()) %>% 
  filter(year >= 2005) %>% 
  ungroup() %>% 
  
  ggplot(aes(x=long, y=lat)) +
  theme_publication() +

  coord_quickmap(xlim=c(-18,2) , ylim=c(54,64)) +
  geom_polygon(data=fao.df, aes(long, lat, group=group), 
               fill = NA, size=0.25, color="gray60", alpha=0.3) +
  geom_polygon(data=world.df, aes(long,lat,group=group), 
               fill="cornsilk", size=0.25,color="gray15", alpha=0.7) +
  geom_point(aes(long, lat, colour = fleet, size=nhauls), alpha=0.5) +
  # scale_fill_brewer(palette = "YlOrRd") + 
  labs(x = NULL, y = NULL) +
  # guides(colour = FALSE) +
  facet_wrap(~year)




```

*`r fig_nums("plothauls")`*

##### page break

**Standardized CPUE index**

We applied the same model for standardization of of CPUE: CPUE ~ year + week + depth, where CPUE is expressed as catch per day and per rectangle. Catches have first been summed by vessel, year, week and rectangle and the number of hauls and fishing days have been calculated. Then the catches and effort (fishing days) have been summed over all vessels by year and week and the average depth has been calculated. CPUE was then calculated as the average catch per rectangle and per day.    

```{r echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}

fig_nums(
  name    = "standardizedcpue", 
  caption = "ARU.27.5b6a standardized CPUE (catch per rectangle and day), in comparison with WKGSS series",
  display = FALSE)

# This is important for the ordering of factors: the base level is set to be
# the last level of the factor. More explanations: ?contrasts
options(contrast=c("contr.SAS","contr.poly"))

dataset <- "comb_arg_byday"
cpuevar <- "cpue_day"

# Specify the model 
form <- as.formula(get(cpuevar) ~ year + week + depth_cat) # this is the model

glm  <- MASS::glm.nb(
  form,                           # input formula
  data=get(dataset),                            # input dataset
  link=log,                             # log link function because of the negative binomial
  na.action=na.exclude,                 # all NAs will be removed
  control=glm.control(epsilon = 1e-08,  # change between iterations; once it is this small it stops
                        maxit = 100 ) # no more than 100 iterations
 )  

# create newdat
newdat <- expand.grid(
  year     = unique(get(dataset)$year),
  week     = unique(get(dataset)$week),
  depth_cat= levels(get(dataset)$depth_cat)) 

# Final GAM model
gam <- mgcv::gam(form, 
           data=get(dataset),
           family=mgcv::negbin(MASS::glm.nb(form,
                                            data=get(dataset), 
                                            link=log,
                                            na.action=na.exclude)$theta))


# get the prediction
pred <- predict(gam,newdata=newdat,se.fit=TRUE,type = "link")


df <-
  make_ci(pred, newdat) %>% 
  mutate(cpue = exp(pred$fit),
         type = paste0("wgdeep", thisyear)) %>% 
  filter(depth_cat == "[300,400)",
         week == "29") %>%   
  dplyr::select(-fit) %>% 
  remove_rownames()

# with(cpue, table(depth_cat, week))

# glimpse(wkgss_cpue_index)

df %>% 
  mutate(year = as.integer(as.character(year))) %>% 
  bind_rows(lastyear_cpue_index) %>% 
  
  ggplot(aes(x=year, y=cpue)) +
  theme_publication() +
  # theme(legend.position = "none") +
  
  # geom_ribbon(data=wkgss_cpue_index, aes(ymin=lwr, ymax=upr, fill=type), alpha=0.2) +
  # geom_path  (data=wkgss_cpue_index, aes(y=cpue, colour=type)) +
  # geom_point (data=wkgss_cpue_index, aes(y=cpue, colour=type)) +
  
  geom_path(aes(y=cpue, colour=type), size=1) +
  geom_point(aes(y=cpue, colour=type), size=2) +
  geom_errorbar(aes(ymin=lwr, ymax=upr, colour=type), width=.2) + 
  expand_limits(y=0) +
  labs(y="cpue (tonnes/day)") 

# plot diagnostics for glm
# par(mfrow=c(2,2)); mgcv::gam.check(glm)

# plot diagnostics for gam
# mgcv::plot.gam(gam, all.terms=T, page=1)
# mgcViz::plot.gamViz(gam, select=1)

# plot diagnostics for glm2
# par(mfrow=c(2,2)); gam.check(gam)

```

*`r fig_nums("standardizedcpue")`*

##### page break

**Model diagnostics**

```{r echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}

fig_nums(
  name    = "diagnostics", 
  caption = "ARU.27.5b6a standardized CPUE model diagnostics",
  display = FALSE)

# plot diagnostics for gam
# mgcv::plot.gam(gam, all.terms=T, page=1)
# mgcViz::plot.gamViz(gam, select=1)

# plot diagnostics for glm2
par(mfrow=c(2,2)); gam.check(gam)

```

*`r fig_nums("diagnostics")`*

**Evaluation of explanatory variables**

```{r echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}

fig_nums(
  name    = "variables", 
  caption = "ARU.27.5b6a standardized CPUE explanatory variables",
  display = FALSE)

# plot diagnostics for gam
mgcv::plot.gam(gam, all.terms=T, page=1)

```

*`r fig_nums("variables")`*

##### page break

```{r echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

tab_nums(
  name    = "cpuetable", 
  caption = "ARU.27.5b6a standardized commercial CPUE (tonnes/day) for greater silversmelt, with lower and upper values based on the standard error.",
  display = FALSE)

df %>%   
  ungroup() %>%
  mutate(year = as.character(year)) %>%
  dplyr::select(year, cpue, lwr, upr) %>%
  pandoc.table(., 
               style        = "simple",
               split.tables = 100, 
               split.cells  = c(rep(7,10)),
               justify      = "right",
               missing      =".",
               big.mark     = ',', 
               round        = c(0,2,2,2,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0))



# save to csv
df %>%   
  mutate(year = as.character(year)) %>% 
  dplyr::select(year, cpue, lwr, upr) %>% 
  write.csv(file=paste0("Standardized commercial cpue 2005-",lastyear,".csv"), row.names = FALSE)

```

*`r tab_nums("cpuetable")`*

##### page break

**Single fleet analysis**

A single fleet analysis was carried out by using the combined raw CPUE datasets and extracting the separate parts for the Faroese and PFA fleets. These data were then processed in a similar fashion as in the combined analysis. It is clear that the Faroese data is substantially more precise that the data from PFA as evident from the confidence intervals. This is likely due to the number of observations, where the dataset from Faroe Islands over all years is based on 10 times the number of hauls compared to the PFA data. In addition, the Faroese fishery is a targetted fisehry for silver smelt, while the PFA fishery is a mixed fishery with blue whiting in the daytime and silver smelt in the nighttime. 

```{r echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}

fig_nums(
  name    = "singlefleet", 
  caption = "ARU.27.5b6a standardized single fleet CPUE (catch per rectangle and day)",
  display = FALSE)

dataset <- "comb_arg_byday"
cpuevar <- "cpue_day"

# Specify the model 
form <- as.formula(get(cpuevar) ~ year + week + depth_cat) # this is the model

# PFA only

cpue_onlypfa     <- 
  filter(get(dataset), fleet == "pfa") %>% group_by(year) %>% filter(year != "2005")

# save(cpue_onlypfa, file="onlypfa.RData")
# cpue_onlypfa %>% group_by() %>% summarize(nhauls=sum(nhauls))

finalMod_onlypfa <- 
  gam(form, data=cpue_onlypfa,
                family=negbin(glm.nb(form,
                                     data=cpue_onlypfa, 
                                     link=log,
                                     na.action=na.exclude)$theta))

newdat_onlypfa <- 
  expand.grid(
    year        = unique(cpue_onlypfa$year),
    week        = unique(cpue_onlypfa$week),
    depth_cat   = unique(cpue_onlypfa$depth_cat))

pred_onlypfa      <- 
  predict(finalMod_onlypfa ,newdat_onlypfa,se.fit=T,type="link")

df_onlypfa <-
  make_ci(pred_onlypfa, newdat_onlypfa) %>% 
  mutate(
    cpue = exp(pred_onlypfa$fit), 
    fleet = "pfa") %>% 
  filter(depth_cat == "[300,400)", week == "21")


# Faroer only

cpue_onlyfar     <- 
  filter(get(dataset), fleet == "faroer" )

finalMod_onlyfar <- 
  gam(form, data=cpue_onlyfar,
                family=negbin(glm.nb(form,
                                     data=cpue_onlyfar,
                                     link=log,
                                     na.action=na.exclude)$theta))

newdat_onlyfar <- expand.grid(
    year        = unique(cpue_onlyfar$year),
    week        = unique(cpue_onlyfar$week),
    depth_cat   = unique(cpue_onlyfar$depth_cat))

# calculate the predicted values and confidence intervals
pred_onlyfar      <- 
  predict(finalMod_onlyfar ,newdat_onlyfar,se.fit=T,type="link")

df_onlyfar <-
  make_ci(pred_onlyfar, newdat_onlyfar) %>% 
  mutate(
    cpue = exp(pred_onlyfar$fit), 
    fleet = "faroer") %>% 
  filter(depth_cat == "[300,400)", week == "21")

bind_rows(df_onlyfar, df_onlypfa) %>% 
  mutate(year = as.numeric(as.character(year))) %>% 
  
  ggplot(aes(year)) +
  theme_publication() +
  theme(legend.position  = "none") +
  geom_ribbon(aes(ymin=lwr, ymax=upr, fill=fleet), alpha=0.2) +
  geom_line(aes(y=cpue, colour=fleet), size=1) +
  geom_point(aes(y=cpue, colour=fleet), size=1.5) +
  expand_limits(y=0) +
  labs(y="cpue (tonnes per day)") +
  facet_wrap(~fleet, ncol=2)

```

*`r fig_nums("singlefleet")`*

# Discussion

CPUE standardization using GLM procedures is a common way of dealing with CPUE information. Here we used aggregated data (catch per day) as the main response variable and year, week and depth category as explanatory variables. by area and period, which cannot use attributes that are related to the hauls. The standardized CPUE for WGDEEP 2021 is highly consistent with the CPUE that was calculated during WGDEEP 2020. 

Both data sources (Faroese data and PFA data) indicate an increase in CPUE in the last 5-6 years although it does not reach the level seen in the late 2000s. The data from the Faroese fisheries are generated from a targetted fishery on silver smelt, while the data from the PFA is from a mixed fishery with blue whiting (blue whiting in the daytime, silver smelt in the nighttime). This probably leads to the higher uncertainties in the CPUE estimates for the PFA compared to the Faroese fleet. It is also noted that the number of observations in the PFA fisheries prior to 2015 is much lower than after 2015, because the self-sampling program only started in 2015. 


# References

Pastoors, M. A. and F. J. Quirijns (2020). Correcting an error in the CPUE Standardizing of Greater silversmelt for WGDEEP 2020, WD05.

Quirijns, F. J. and M. A. Pastoors (2020). CPUE standardization for greater silversmelt in 5b6a. WKGSS 2020, WD03.
