---
output: 
  word_document:
    reference_docx: ../report_template_v1.6.dotx
---

```{r setup, include=FALSE}
# -------------------------------------------------------------------------------------
# CPUE standardizing for argentines: pfa + faroer data
#
# 23/04/2020 modified the calculation of CPUE (catch by day and by rectangle)
# 25/04/2020 added option for standardization by haul; rearranged data processing steps
# -------------------------------------------------------------------------------------

# ================================================================================
# 0. Initialization. Set working directory and filename and sourcing
# ================================================================================

require("knitr")
knitr::opts_chunk$set(echo=FALSE,fig.width=10,message=FALSE,warning=FALSE,comment="",crop=TRUE)
knit_hooks$set(crop = hook_pdfcrop)

rm(list=ls())

global_plot_map   <- TRUE
global_plot_fao   <- TRUE
global_plot_depth <- FALSE

# Libraries
library(rmarkdown)
library(maps)          # world maps
library(sp)            # spatial data 
require(geosphere, quietly=TRUE)           # spatial manipulation

library(pander)        # tables
library(colorspace)    # hcl colours
library(viridis)       # Viridis colour scheme
library(RColorBrewer)  # ColourBrewer

library(tidyverse)     # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
library(lubridate)     # data handling
library(reshape2)      # reshaping data; e.g. cast
library(readxl)        # excel reader
library(broom)         # clean up statistics
library(scales)        # pretty scales
library(stringr)       # string manipulations
library(magrittr)      # for e.g. set_colnames
library(captioner)     # captioning of figures and tables

library(mgcv)          # tensor spline for gam
library(lme4)
library(MASS)
library(mgcViz)         # GAM output to ggplot

# To number figures and tables
fig_nums <- captioner::captioner(prefix = "Figure ")
tab_nums <- captioner::captioner(prefix = "Table ")

# default settings for tables
panderOptions('table.alignment.default', function(df) ifelse(sapply(df, is.numeric), 'right', 'left'))

# set onedrive directory
onedrive <- file.path(Sys.getenv('USERPROFILE'), 'PFA/PFA team site - PRF') 

# set data path
wgdeeppath <- "//community.ices.dk/ExpertGroups/WGDEEP/2020 Meeting Docs/06. Data/aru.27.5b6a/CPUE"
wkgsspath  <- "//community.ices.dk/ExpertGroups/benchmarks/2020/wkdeep/2014 Meeting docs/06. Data/aru.27.5b6a/CPUE"
  
# load spatial datasets
load(file.path(onedrive, "rdata/world.df.RData"))
load(file.path(onedrive, "rdata/fao.df.RData"))
load(file.path(onedrive, "rdata/depth200.df.RData"))
load(file.path(onedrive, "rdata/fao.RData"))

rect <- 
  get(load(file.path(onedrive, "rdata/icesrectangles.df.RData"))) %>% 
  group_by(rect) %>% 
  filter(row_number() == 1) %>% 
  dplyr::select(rect, long, lat)

# Source the lowcase function
source("lowcase.r")
source("theme_publication.r")

# CI function
make_ci <- function(pred, data){
  fit <- pred$fit
  lwr <- exp(pred$fit - (1.96 * pred$se.fit))
  upr <- exp(pred$fit + (1.96 * pred$se.fit))
  return(data.frame(fit, lwr, upr, data))
  }

# --------------------------------------------------------------------
# read in information as used in WKGSS 2020
# --------------------------------------------------------------------

faroer_aru_raw_wkgss <- 
  read_excel(file.path(wkgsspath, "Faroe GSS_CommercialTrawlData1995-2019_WKGSS.xlsx"), col_names = TRUE, col_types = "text") %>% 
  lowcase() %>% 
  setNames(gsub("icesrect","rect", names(.))) %>% 
  setNames(gsub("hours","duration", names(.))) %>% 
  setNames(gsub("gsskg","catch", names(.))) %>% 
  setNames(gsub("depthm","depth", names(.))) %>% 
  
  mutate_at(c("year", "day", "vessel"), list(as.integer)) %>% 
  mutate_at(c("duration", "catch", "depth"), list(as.numeric)) %>% 
  
  # only data after 2005 (decision: Lise Ofstad, because since then >80% of the catch is covered)
  filter (year > 2004) %>% 
  
  mutate(
    catch = catch/1000,   # catch in tonnes
    logcatch = log(catch), 
    date = as.Date(day, origin="1899-12-30"),
    day  = yday(date),
    week = as.numeric(week(date)),
    time = format(strptime(sprintf("%04d", as.numeric(shoottime)), format="%H%M"), 
                       format = "%H:%M"),
    shoottime = ymd_hm(paste(date, time))) %>% 
  
  left_join(rect, by = "rect") %>% 
  mutate(fleet="faroer") %>% 
  dplyr::select(-time, -status, -gsscpue, -haulkg, -rect, -faroearea, -logcpue) 

# arg <- get(load(file.path(onedrive, "rdata/arg.RData")))
# arg$vesselid <- arg %>%  group_indices(vessel)
# arg <- arg %>% dplyr::select(-vessel) %>% rename(vessel = vesselid)
# save(arg, file=file.path(onedrive, "rdata/arg.RData"))  

# PFA raw CPUE at WKGSS
pfa_arg_raw_wkgss <-
  get(load(file.path(wkgsspath, "PFA arg.RData"))) %>% 
  setNames(gsub("calc_depth","depth", names(.))) %>% 
  setNames(gsub("shoot_time","shoottime", names(.))) %>% 
  
  filter (year > 2004,   # only data after 2005 (as for Faroese data)
          division %in% c("27.5.b", "27.6.a")) %>% 
  
  mutate(species = ifelse(species %in% c("aru","ary","arg"), "arg","oth")) %>% 
  filter(species == "arg") %>% 
  group_by(vessel, trip, haul, shoottime, depth, date, year, rect) %>%
  summarize(catch = sum(catch, na.rm=TRUE), 
            duration = mean(duration, na.rm=TRUE)) %>% 
  mutate(
    logcatch = log(catch),
    week = as.numeric(week(date)),
    month= as.factor(month(date))) %>% 
  
  ungroup() %>% 

  left_join(rect, by = "rect") %>% 
  mutate(fleet = "pfa") %>% 
  filter(!(vessel=="SCH302" & trip=="2017E"))  %>%   # trip with incorrect start and end times
  dplyr::select(-rect, -trip, -haul)


# combined cpue at WKGSS
comb_arg_byday_wkgss <- 
  bind_rows(faroer_aru_raw_wkgss, pfa_arg_raw_wkgss) %>% 
  ungroup() %>% 
  mutate (month = as.factor(month),
          day = day(date),
          hset = as.numeric(strftime(shoottime, format="%H"))) %>% 
  mutate (depth_cat= cut(depth, breaks=c(0, 200, 300, 400, 500, 600, 700, 2500), right = FALSE),
          long_cat = as.factor(long),
          lat_cat  = as.factor(lat),
          hset_cat = cut(hset, include.lowest = TRUE, breaks=c(0,2,4,6,8,10,12,14,16,18,20,22,24))) %>% 
  filter(duration < 12,    # sometimes duration > 20 hours. Issue with datetimes to be solved
         long > -18,
         long < 5,
         !is.na(depth)) %>%
  
  group_by(year,month,week,depth_cat,lat_cat,long_cat,hset_cat,fleet) %>% 
  summarise(catch    = sum(catch, na.rm=TRUE),
            nhauls   = n(),
            nvessels = n_distinct(vessel),
            duration = sum(duration, na.rm=TRUE),
            ndays    = n_distinct(day, na.rm = FALSE)
            ) %>% 
  mutate(
    effort = ndays,
    cpue = catch/effort,
    logcatch = log(catch),
    lcpue= logcatch/effort)  %>%
  ungroup() %>% 
  filter(week %in% c(15:34) ) %>% 
  filter(lat_cat %in% (c("59.5", "60", "60.5", "61", "61.5", "62", "62.5"))) %>% 
  
  # explanatory variables should be factors
  mutate(year = as.factor(year),
         week = as.factor(week)) %>% 
  dplyr::select(-month,-logcatch,-lcpue, -nvessels) 

# WKGSS 2020 CPUE index
wkgss_cpue_index <-
  read.csv(file.path(wkgsspath, "ARU.27.5b6a standardized CPUE.csv")) %>% 
  mutate(type="wkgss")

# --------------------------------------------------------------------
# read in information as used in WGDEEP 2020
# --------------------------------------------------------------------

#  Load the PFA CPUE data

# source("../../prf/R/PFA make ARU datasets.r")
pfa_arg_raw <- 
  get(load(file.path(wgdeeppath, "pfa_arg_raw.RData"))) %>% mutate(fleet="pfa") 
# summarize(pfa_arg_raw, nhauls=n())

# Load the Faroer CPUE data

faroer_aru_raw <-
    read_excel(file.path(wgdeeppath, "Faroe GSS_CommercialTrawlData1995-2019_WKGSS.xlsx"), 
             col_names = TRUE,
             col_types = "text") %>% 
  lowcase() %>% 
  setNames(gsub("icesrect","rect", names(.))) %>% 
  setNames(gsub("hours","duration", names(.))) %>% 
  setNames(gsub("gsskg","catch", names(.))) %>% 
  setNames(gsub("depthm","depth", names(.))) %>% 
  
  mutate_at(c("vessel", "year", "month", "day"), list(as.integer)) %>% 
  mutate_at(c("duration", "catch", "depth"), list(as.numeric)) %>% 

  mutate(
    catch = catch/1000,   # catch in tonnes
    date  = as.Date(day, origin="1899-12-30"),
    day   = day(date),
    yday  = yday(date),
    week  = as.numeric(week(date)),
    time  = format(strptime(sprintf("%04d", as.numeric(shoottime)), format="%H%M"), 
                   format = "%H:%M"),
    shoot_time = ymd_hm(paste(date, time)), 
    fleet = "faroer", species="aru") %>% 
  
  ungroup() %>% 
  dplyr::select(vesselid=vessel, date, year, month, week, day, yday, shoot_time, depth,
                catch, duration, rect, species, fleet)

comb_arg_byhaul <-
  bind_rows(faroer_aru_raw, pfa_arg_raw) %>% 
  
  # add rectangle information
  left_join(rect, by = "rect") %>% 
  
  # add hset time and depth category 
  mutate(hset      = as.numeric(strftime(shoot_time, format="%H"))) %>% 
  mutate(hset_cat  = cut(hset, include.lowest = TRUE, breaks=c(0,2,4,6,8,10,12,14,16,18,20,22,24))) %>% 
  mutate(depth_cat = cut(depth, breaks=c(0, 200, 300, 400, 500, 600, 700, 2500), right = FALSE)) %>% 
  
  mutate(cpue_hour = catch / duration) %>% 
  mutate(cpue_haul = catch) %>% 
  
  # filter area and weeks
  filter(long > -18, long < 5) %>% 
  # filter(lat  >=59.5, lat <= 62.5) %>% 
  filter(week %in% c(15:34)) %>% 
  filter(year > 2004) %>% 
  filter(!is.na(depth)) 

# comb_arg_byhaul %>% filter(fleet=="pfa") %>% View()
# pfa_arg_raw %>% left_join(rect) %>% dplyr::select(lat) %>% hist(as.numeric(lat))
# comb_arg_byhaul %>%  filter(fleet=="pfa") %>% ggplot(aes(lat)) + geom_histogram()

comb_arg_byday <-
  comb_arg_byhaul %>% 
  
  # aggregate within vessels and days
  group_by(fleet, vesselid, year, week, rect, long, lat) %>%
  summarize(catch    = sum(catch, na.rm=TRUE), 
            duration = sum(duration, na.rm=TRUE),
            nhauls   = n(),
            ndays    = n_distinct(date),
            depth    = mean(depth, na.rm=TRUE)) %>% 
  arrange(fleet, year, week, vesselid) %>% 
  # filter(fleet=="pfa")

  # aggregate within vessels and days
  # group_by(fleet, vesselid, year, month, week, day, rect, long, lat) %>%
  # arrange(fleet, year, week, vesselid) %>% 
  # summarize(catch    = sum(catch, na.rm=TRUE), 
  #           duration = sum(duration, na.rm=TRUE),
  #           nhauls   = n(),
  #           ndays    = 1,
  #           depth    = mean(depth, na.rm=TRUE)) %>% 
  
  # sum over vessels and days
  group_by(fleet, year, week, rect, long, lat) %>%
  summarize(catch    = sum(catch, na.rm=TRUE), 
            duration = sum(duration, na.rm=TRUE),
            nhauls   = sum(nhauls, na.rm=TRUE),
            ndays    = sum(ndays, na.rm=TRUE),
            depth    = mean(depth, na.rm=TRUE)) %>% 
  
  # calculate CPUE
  mutate(cpue_day = catch / ndays) %>% 
  mutate(depth_cat= cut(depth, breaks=c(0, 200, 300, 400, 500, 600, 700, 2500), right = FALSE)) %>% 
  
  ungroup() %>% 
  
  # explanatory variables should be factors
  mutate(year = as.factor(year),
         week = as.factor(week),
         long_cat = as.factor(long),
         lat_cat = as.factor(lat)) 

# comb_arg_byday %>% filter(fleet=="pfa") %>% View()

# Overview of number of hauls
# comb_arg_byday %>%
#   group_by(fleet, year) %>%
#   summarize(nhauls = sum(nhauls, na.rm=TRUE)) %>%
#   ggplot(aes(x=as.numeric(as.character(year)), y=nhauls, group=fleet)) +
#   geom_line(aes(colour=fleet))

# comb_arg_byday %>%
#   group_by(fleet, year) %>%
#   summarize(catch = sum(catch, na.rm=TRUE)) %>%
#   ggplot(aes(x=as.numeric(as.character(year)), y=catch, group=fleet)) +
#   geom_line(aes(colour=fleet))


```


**Correcting an error in the CPUE Standardizing of Silversmelt for WGDEEP 2020**

M.A. Pastoors & F.J. Quirijns

Corresponding author: mpastoors@pelagicfish.eu

`r format(Sys.time(), '%d/%m/%Y')`

**Abstract ** 

At the WKGSS 2020 benchmark of Greater silver smelt in 5b and 6a, a combined and standardized CPUE series for the Faroe and EU fleets has been introduced (Quirijns and Pastoors 2020). On checking the data in preparation for WGDEEP 2020, a small error was detected in the way CPUE was calculated. This report provides a summary of the issue and proposed a method to repair the situation. The overall trend in CPUE is still similar although there are some differences in the most recent year. 


<!--1. Introduction ------------------------------------------------------ -->

# Introduction

At the WKGSS 2020 benchmark of Greater silver smelt in 5b and 6a, a combined and standardized CPUE series for the Faroe and EU fleets has been introduced (Quirijns and Pastoors 2020). On checking the data in preparation for WGDEEP 2020, a small error was detected in the way CPUE was calculated. This report provides a summary of the issue and proposed a method to repair the situation. 

# Issues

The CPUE standardization was based on a GLM model with CPUE against year, week and depth category. The calculation of CPUE was intended to be based on catch per day and per rectangle which was supposed to be achieved by summing the catches of all vessels in a certain rectangle, calculating the combined number of days spent by the vessels in the rectangle and dividing the catch by the number of days for that rectangle. 

On checking the code to prepare for WGDEEP 2020, we discovered two issues in the code: 1) the calculation of CPUE and 2) the selection of hauls to be included. 

1) Calculation of CPUE. Because we had initially planned to test different potential explanatory variables in the GLM model, we had included the hour for shooting the haul (hset) as a factor. This factor did not contribute significantly to the GLM and was left out in the final calculation. However, because the hset variable was included when calculating the catch per day and per rectangle, effectively, we were not calculating the catch per day, but rather the catch per haul, as the hset variable meant that few vessels would have the same hset when fishing in the same rectangle on the same day.

2) Selection of hauls to be included. We discovered that only hauls were included north of 59.5 degrees latitude, as this was the area covered by the Faroe fleet. However, the PFA fleet carries out the main fishery south of 59.6 degrees. Due to the spatial selection process almost no hauls of the PFA fleet were included in the final calculation. 

# Results

The first issue is illustrated with the effort metrics used in the calculation for WKGSS 2020 and the new calculation for WGDEEP 2020 (figure 1). 

```{r echo=FALSE, fig.asp=0.5, fig.align="center", message=FALSE, warning=FALSE}

fig_nums(
  name    = "effort", 
  caption = "ARU.27.5b6a Effort (fishing days per rectangle) used for CPUE calculation",
  display = FALSE)

bind_rows(
  comb_arg_byday_wkgss %>% mutate(source="wkgss") %>% dplyr::select(year, effort, source),
  comb_arg_byday       %>% mutate(source="wgdeep") %>% dplyr::select(year, effort=ndays, source),
) %>% 
  mutate(
    year = as.integer(as.character(year)),
    source = factor(source, levels=c("wkgss","wgdeep"))) %>% 
  
  ggplot(aes(x=year, y=effort, group=source)) +
  theme_publication() +
  theme(legend.position="none") +
  geom_boxplot(aes(colour=source, group=year)) +
  facet_wrap(~source)


```
*`r fig_nums("effort")`*

The 'raw' CPUE, based on the effort metrics in figure 1 is then calculated as shown below, whereby the WKGSS effort is effectively based on the catch per haul and the new effort of the catch per day. 

```{r echo=FALSE, fig.asp=0.5, fig.align="center", message=FALSE, warning=FALSE}

fig_nums(
  name    = "cpue", 
  caption = "ARU.27.5b6a Catch per unit effort.",
  display = FALSE)

bind_rows(
  comb_arg_byday_wkgss %>% mutate(source="wkgss") %>% dplyr::select(year, cpue, source),
  comb_arg_byday       %>% mutate(source="wgdeep") %>% dplyr::select(year, cpue=cpue_day, source),
) %>% 
  mutate(
    year = as.integer(as.character(year)),
    source = factor(source, levels=c("wkgss","wgdeep"))) %>% 
  
  ggplot(aes(x=year, y=cpue, group=source)) +
  theme_publication() +
  theme(legend.position="none") +
  geom_boxplot(aes(colour=source, group=year)) +
  facet_wrap(~source)


```

*`r fig_nums("cpue")`*

##### page break

The second issue (selection of hauls to be included) is illustrated by the number of hauls by fleet and year in the figure below. 

```{r echo=FALSE, fig.asp=0.5, fig.align="center", message=FALSE, warning=FALSE}

fig_nums(
  name    = "nhauls", 
  caption = "ARU.27.5b6a number of hauls used for the CPUE calculation.",
  display = FALSE)

t <- bind_rows(
  comb_arg_byday_wkgss %>% mutate(source="wkgss") %>% 
    group_by(source, fleet, year) %>% summarize(nhauls = sum(nhauls, na.rm=TRUE)), 
  comb_arg_byday       %>% mutate(source="wgdeep") %>% 
    group_by(source, fleet, year) %>% summarize(nhauls = sum(nhauls, na.rm=TRUE))
) %>% 
  ungroup() %>% 
  mutate(
    year = as.integer(as.character(year)),
    source = factor(source, levels=c("wkgss","wgdeep")))

t1 <-
  t %>% 
  group_by(source, fleet) %>% 
  summarize(nhauls = sum(nhauls, na.rm=TRUE)) %>% 
  group_by(source) %>% 
  summarize(string = paste(fleet,nhauls, collapse=", "))

t %>%   
  ggplot(aes(x=year, y=nhauls)) +
  theme_publication() +
  theme(legend.position="none") +
  geom_line(aes(colour=fleet, group=fleet)) +
  # annotate(geom = 'text', aes(label = 'sometext', x = -Inf, y = Inf), hjust = 0, vjust = 1) +
  geom_label(data=t1, aes(x=-Inf, y=Inf,label=string), inherit.aes = FALSE, hjust=0, vjust=1) +
  facet_wrap(~source)


```

*`r fig_nums("nhauls")`*

For the years 2015-2019, below are the spatial distributions of the used number of hauls by fleet. 

```{r echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}

fig_nums(
  name    = "plothauls", 
  caption = "ARU.27.5b6a plot of the number of hauls by rectangle and day",
  display = FALSE)

t1 <- 
  comb_arg_byday_wkgss %>% 
  mutate(
    year = as.numeric(as.character(year)),
    lat  = as.numeric(as.character(lat_cat)),
    long = as.numeric(as.character(long_cat)),
    source="wkgss") %>% 
  group_by(source, fleet, year, lat, long) %>% 
  summarize(nhauls = sum(nhauls, na.rm=TRUE))

t2 <- 
  bind_rows(pfa_arg_raw, faroer_aru_raw) %>% 
  left_join(rect) %>% 
  mutate(
    year = as.numeric(as.character(year)),
    source="wgdeep") %>% 
  group_by(source, fleet, year, lat, long) %>% 
  summarize(nhauls = n())

bind_rows(t1,t2) %>%
  filter(year >= 2015) %>% 
  ungroup() %>% 
  mutate(source = factor(source, levels=c("wkgss","wgdeep"))) %>% 
  ggplot(aes(x=long, y=lat)) +
  theme_publication() +

  coord_quickmap(xlim=c(-18,2) , ylim=c(54,64)) +
  geom_polygon(data=fao.df, aes(long, lat, group=group), 
               fill = NA, size=0.25, color="gray60", alpha=0.3) +
  geom_polygon(data=world.df, aes(long,lat,group=group), 
               fill="cornsilk", size=0.25,color="gray15", alpha=0.7) +
  geom_point(aes(long, lat, colour = fleet, size=nhauls), alpha=0.5) +
  # scale_fill_brewer(palette = "YlOrRd") + 
  labs(x = NULL, y = NULL) +
  facet_grid(source~year)




```

*`r fig_nums("plothauls")`*


**New standardized CPUE index**

We applied the same model for standardization of of CPUE: CPUE ~ year + week + depth, where CPUE is now actually expressed as catch per day and per rectangle. Catches have first been summed by vessel, year, week and rectangle and the number of hauls and fishing days have been calculated. Then the catches and effort (fishing days) have been summed over all vessels by year and week and the average depth has been calculated. CPUE was then calculated as the average catch per rectangle and per day. In addition we removed the constraint on only using hauls north of 59.5 degrees latitude. This follows the intended procedure explained in Quirijns and Pastoors (2020).   

```{r echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}

fig_nums(
  name    = "standardizedcpue", 
  caption = "ARU.27.5b6a standardized CPUE (catch per rectangle and day), in comparison with WKGSS series",
  display = FALSE)

# This is important for the ordering of factors: the base level is set to be
# the last level of the factor. More explanations: ?contrasts
options(contrast=c("contr.SAS","contr.poly"))

dataset <- "comb_arg_byday"
cpuevar <- "cpue_day"

# Specify the model 
form <- as.formula(get(cpuevar) ~ year + week + depth_cat) # this is the model

glm  <- MASS::glm.nb(
  form,                           # input formula
  data=get(dataset),                            # input dataset
  link=log,                             # log link function because of the negative binomial
  na.action=na.exclude,                 # all NAs will be removed
  control=glm.control(epsilon = 1e-08,  # change between iterations; once it is this small it stops
                        maxit = 100 ) # no more than 100 iterations
 )  

# create newdat
newdat <- expand.grid(
  year     = unique(get(dataset)$year),
  week     = unique(get(dataset)$week),
  depth_cat= levels(get(dataset)$depth_cat)) 

# Final GAM model
gam <- mgcv::gam(form, 
           data=get(dataset),
           family=mgcv::negbin(MASS::glm.nb(form,
                                            data=get(dataset), 
                                            link=log,
                                            na.action=na.exclude)$theta))


# get the prediction
pred <- predict(gam,newdata=newdat,se.fit=TRUE,type = "link")


df <-
  make_ci(pred, newdat) %>% 
  mutate(cpue = exp(pred$fit),
         type = "wgdeep") %>% 
  filter(depth_cat == "[300,400)",
         week == "29") %>%   
  dplyr::select(-fit)

# with(cpue, table(depth_cat, week))

# glimpse(wkgss_cpue_index)

df %>% 
  mutate(year = as.integer(as.character(year))) %>% 
  bind_rows(wkgss_cpue_index) %>% 
  
  ggplot(aes(x=year, y=cpue)) +
  theme_publication() +
  # theme(legend.position = "none") +
  
  # geom_ribbon(data=wkgss_cpue_index, aes(ymin=lwr, ymax=upr, fill=type), alpha=0.2) +
  # geom_path  (data=wkgss_cpue_index, aes(y=cpue, colour=type)) +
  # geom_point (data=wkgss_cpue_index, aes(y=cpue, colour=type)) +
  
  geom_path(aes(y=cpue, colour=type), size=1) +
  geom_point(aes(y=cpue, colour=type), size=2) +
  geom_errorbar(aes(ymin=lwr, ymax=upr, colour=type), width=.2) + 
  expand_limits(y=0) +
  labs(y="cpue (tonnes/day)") 

# plot diagnostics for glm
# par(mfrow=c(2,2)); mgcv::gam.check(glm)

# plot diagnostics for gam
# mgcv::plot.gam(gam, all.terms=T, page=1)
# mgcViz::plot.gamViz(gam, select=1)

# plot diagnostics for glm2
# par(mfrow=c(2,2)); gam.check(gam)

```

*`r fig_nums("standardizedcpue")`*

##### page break

Model diagnostics

```{r echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}

fig_nums(
  name    = "diagnostics", 
  caption = "ARU.27.5b6a standardized CPUE model diagnostics",
  display = FALSE)

# plot diagnostics for gam
# mgcv::plot.gam(gam, all.terms=T, page=1)
# mgcViz::plot.gamViz(gam, select=1)

# plot diagnostics for glm2
par(mfrow=c(2,2)); gam.check(gam)

```

*`r fig_nums("diagnostics")`*

Evaluation of explanatory variables

```{r echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}

fig_nums(
  name    = "variables", 
  caption = "ARU.27.5b6a standardized CPUE explanatory variables",
  display = FALSE)

# plot diagnostics for gam
mgcv::plot.gam(gam, all.terms=T, page=1)

```

*`r fig_nums("variables")`*

```{r echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

tab_nums(
  name    = "cpuetable", 
  caption = "ARU.27.5b6a standardized commercial CPUE (tonnes/day) for greater silversmelt, with lower and upper values based on the standard error.",
  display = FALSE)

df %>%   
  mutate(year = as.character(year)) %>% 
  dplyr::select(year, cpue, lwr, upr) %>% 
  pandoc.table(., 
               style        = "simple",
               split.tables = 100, 
               split.cells  = c(rep(7,10)),
               justify      = "right",
               missing      =".",
               big.mark     = ',', 
               round        = c(0,2,2,2,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0))



# save to csv
df %>%   
  mutate(year = as.character(year)) %>% 
  dplyr::select(year, cpue, lwr, upr) %>% 
  write.csv(file=paste0("Standardized commercial cpue 2005-2019.csv"), row.names = FALSE)

```

*`r tab_nums("cpuetable")`*

##### page break

**Single fleet analysis**

A single fleet analysis was carried out by using the combined raw CPUE datasets and extracting the separate parts for the Faroese and PFA fleets. These data were then processed in a similar fashion as in the combined analysis. It is clear that the Faroese data is substantially more precise that the data from PFA as evident from the confidence intervals. This is likely due to the number of observations, where the dataset from Faroe Islands over all years is based on 10 times the number of hauls compared to the PFA data. 

```{r echo=FALSE, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE}

fig_nums(
  name    = "singlefleet", 
  caption = "ARU.27.5b6a standardized single fleet CPUE (catch per rectangle and day)",
  display = FALSE)

dataset <- "comb_arg_byday"
cpuevar <- "cpue_day"

# Specify the model 
form <- as.formula(get(cpuevar) ~ year + week + depth_cat) # this is the model

# PFA only

cpue_onlypfa     <- 
  filter(get(dataset), fleet == "pfa") %>% group_by(year) %>% filter(year != "2005")

# save(cpue_onlypfa, file="onlypfa.RData")
# cpue_onlypfa %>% group_by() %>% summarize(nhauls=sum(nhauls))

finalMod_onlypfa <- 
  gam(form, data=cpue_onlypfa,
                family=negbin(glm.nb(form,
                                     data=cpue_onlypfa, 
                                     link=log,
                                     na.action=na.exclude)$theta))

newdat_onlypfa <- 
  expand.grid(
    year        = unique(cpue_onlypfa$year),
    week        = unique(cpue_onlypfa$week),
    depth_cat   = unique(cpue_onlypfa$depth_cat))

pred_onlypfa      <- 
  predict(finalMod_onlypfa ,newdat_onlypfa,se.fit=T,type="link")

df_onlypfa <-
  make_ci(pred_onlypfa, newdat_onlypfa) %>% 
  mutate(
    cpue = exp(pred_onlypfa$fit), 
    fleet = "pfa") %>% 
  filter(depth_cat == "[300,400)", week == "21")


# Faroer only

cpue_onlyfar     <- 
  filter(get(dataset), fleet == "faroer" )

finalMod_onlyfar <- 
  gam(form, data=cpue_onlyfar,
                family=negbin(glm.nb(form,
                                     data=cpue_onlyfar,
                                     link=log,
                                     na.action=na.exclude)$theta))

newdat_onlyfar <- expand.grid(
    year        = unique(cpue_onlyfar$year),
    week        = unique(cpue_onlyfar$week),
    depth_cat   = unique(cpue_onlyfar$depth_cat))

# calculate the predicted values and confidence intervals
pred_onlyfar      <- 
  predict(finalMod_onlyfar ,newdat_onlyfar,se.fit=T,type="link")

df_onlyfar <-
  make_ci(pred_onlyfar, newdat_onlyfar) %>% 
  mutate(
    cpue = exp(pred_onlyfar$fit), 
    fleet = "faroer") %>% 
  filter(depth_cat == "[300,400)", week == "21")

bind_rows(df_onlyfar, df_onlypfa) %>% 
  mutate(year = as.numeric(as.character(year))) %>% 
  
  ggplot(aes(year)) +
  theme_publication() +
  theme(legend.position  = "none") +
  geom_ribbon(aes(ymin=lwr, ymax=upr, fill=fleet), alpha=0.2) +
  geom_line(aes(y=cpue, colour=fleet), size=1) +
  geom_point(aes(y=cpue, colour=fleet), size=1.5) +
  expand_limits(y=0) +
  labs(y="cpue (tonnes per day)") +
  facet_wrap(~fleet, ncol=2)

```

*`r fig_nums("singlefleet")`*

# Discussion

CPUE standardization using GLM procedures is a common way of dealing with CPUE information. We found that the data processing, prior to using it in the CPUE model, is perhaps a somewhat overlooked attribute of standardization. This is especially the case when the data formats of contributing data sets are in different shares and resolutions. Here we concluded that the checking how the subsetting of data is working, should have been better checked. 

When using GLM modelling, a choice needs to be made between using the haul by haul information, which could use attributes related to the haul operation, or using aggregated data by area and period, which cannot use attributes that are related to the hauls. 

Fortunately, the differences in CPUE trends between the analysis for WKGSS 2020 and WGDEEP 2020 are not very large so that the impact on the assessment is expected to be small. 


# References

Quirijns, F. J. and M. A. Pastoors (2020). CPUE standardization for greater silversmelt in 5b6a. WKGSS 2020, WD03.

